---
title: "Explore annomalies"
author: "AJ Perez-Luque (@ajpelu)"
date: "2016 Aug"
output:  
  md_document:
    variant: markdown_github
---
```{r, echo=FALSE, message=FALSE}
require(knitr)
opts_chunk$set(fig.align='center', fig.width = 14, echo=FALSE, message = FALSE, warning = FALSE) 
```

```{r wd, echo=FALSE}
#---------------------------------
machine <- 'ajpelu'
#machine <- 'ajpeluLap'
di <- paste('/Users/', machine, '/Dropbox/phd/phd_repos/qpyr_resilience', sep='')
#---------------------------------
```


```{r packages, warning=FALSE, message=FALSE}
library("dplyr")
library("ggplot2")
library("RCurl")
```


```{r}
# Read and prepare data
anomalo_sn <- read.csv(text = getURL("https://raw.githubusercontent.com/ajpelu/qpyr_resilience/master/data/anomalies/anomalo_sn.csv"), header=T) 
anomalo_clu <- read.csv(text = getURL("https://raw.githubusercontent.com/ajpelu/qpyr_resilience/master/data/anomalies/anomalo_clu.csv"), header=T) 
anomalo_pop <- read.csv(text = getURL("https://raw.githubusercontent.com/ajpelu/qpyr_resilience/master/data/anomalies/anomalo_pop.csv"), header=T) 

std_a_2005_sn <- read.csv(text = getURL("https://raw.githubusercontent.com/ajpelu/qpyr_resilience/master/data/anomalies/std_a_2005_sn.csv"), header=T)
std_a_2012_sn <- read.csv(text = getURL("https://raw.githubusercontent.com/ajpelu/qpyr_resilience/master/data/anomalies/std_a_2012_sn.csv"), header=T)

std_a_2005_clu <- read.csv(text = getURL("https://raw.githubusercontent.com/ajpelu/qpyr_resilience/master/data/anomalies/std_a_2005_clu.csv"), header=T)
std_a_2012_clu <- read.csv(text = getURL("https://raw.githubusercontent.com/ajpelu/qpyr_resilience/master/data/anomalies/std_a_2012_clu.csv"), header=T)

std_a_2005_pop <- read.csv(text = getURL("https://raw.githubusercontent.com/ajpelu/qpyr_resilience/master/data/anomalies/std_a_2005_pop.csv"), header=T)
std_a_2012_pop <- read.csv(text = getURL("https://raw.githubusercontent.com/ajpelu/qpyr_resilience/master/data/anomalies/std_a_2012_pop.csv"), header=T)
```



```{r}
label_variable <- c('myevisp' = 'EVI', 'myndvisp' = 'NDVI')  

label_pop <- c('1' = 'Lugros', '2' = 'Guejar-Sierra', '3' = 'Monachil',
                      '4' = 'Dilar', '5' = 'Durcal', '6' = 'Caniar', 
                      '7' = 'Poqueira', '8' = 'Trevelez')
```

## Annomalies for all *Q. pyrenaica* forests
### EVI

```{r} 
# EVI for all populations 
myylab <- 'EVI'
anomalo_sn %>% 
  dplyr::filter(variable == 'myevisp') %>% 
  ggplot(aes(x=as.factor(month), y=mean, group=event, colour=event)) + 
  geom_point(size=3) + geom_line() + theme_bw() + 
  geom_errorbar(aes(ymin =mean - se, ymax = mean + se, group=event, colour=event), width=0.45) +
  xlab('Months') + ylab(myylab) + 
  # facet_wrap(~variable, 
  #            labeller = as_labeller(label_variable),
  #            scales = 'free_y') +
  theme(strip.background = element_rect(fill = "white"))
```

### NDVI
```{r} 
# EVI for all populations 
myylab <- 'NDVI'
anomalo_sn %>% 
  dplyr::filter(variable == 'myndvisp') 
%>% 
  ggplot(aes(x=as.factor(month), y=mean, group=event, colour=event)) + 
  geom_point(size=3) + geom_line() + theme_bw() + 
  geom_errorbar(aes(ymin =mean - se, ymax = mean + se, group=event, colour=event), width=0.45) +
  xlab('Months') + ylab('myylab') + 
  # facet_wrap(~variable, 
  #            labeller = as_labeller(label_variable),
  #            scales = 'free_y') +
  theme(strip.background = element_rect(fill = "white"))
```


```{r}


myylab <- 'EVI'
anomalo_sn %>% 
  # dplyr::filter(variable == 'myevisp') %>% 
  ggplot(aes(x=as.factor(month), y=mean, group=event, colour=event)) + 
  geom_point(size=3) + geom_line() + theme_bw() + 
  geom_errorbar(aes(ymin =mean - se, ymax = mean + se, group=event, colour=event), width=0.45) +
  xlab('Months') + ylab(myylab) + 
  # facet_wrap(~variable, 
  #            labeller = as_labeller(label_variable),
  #            scales = 'free_y') +
  theme(strip.background = element_rect(fill = "white"))





# Note 

computeAnomaly <- function(df, disturb_years){
  require('dplyr')
  
  rd <- df %>% 
    # Factor Correction to evi and ndvi
    mutate(myevi = evi * 0.0001,
           myndvi = evi * 0.0001) %>% 
    select(-evi) %>% 
    select(-ndvi) %>% 
    
    # Compute year and month 
    mutate(year = lubridate::year(date),  
           month = lubridate::month(date)) %>% 
    
    # Add variable of population cluster and remove population '9'
    mutate(clu_pop = as.factor(ifelse(poblacion == 1, 'Camarate',
                                      ifelse(poblacion %in% c(2,3,4,5), 'Northern slope',
                                             ifelse(poblacion %in% c(6,7,8), 'Southern slope', 'out'))))) %>% 
    filter(clu_pop != 'out') %>% 
    
    # Categorize reference period and disturbance year
    mutate(event = ifelse(year != disturb_years,
                          'ref',
                          ifelse(year == disturb_years[1], 
                                 as.character(disturb_years[1]),
                                 as.character(disturb_years[2])))) %>% 
    
    # Spatial aggregation. All pixels of the same populations are aggregate by mean
    group_by(clu_pop, poblacion, year, month, event) %>% 
    summarise(myevisp = mean(myevi), 
              myndvisp = mean(myndvi)) 
  }


d <- computeAnomaly(df=rawdata, disturb_year = c(2005, 2012))
# d2 <- computeAnomaly(df=rawdata, disturb_year = 2012)
  


### Loop for temporal aggregation 

# misdataframes <- c("d1", "d2")
misdataframes <- c("d")
variables <- c('myevisp', 'myndvisp')

for (j in misdataframes){ 
  
  mydf <- get(j)
  
  # Create three empty dataframe
  anomalo_clu <- data.frame() 
  anomalo_pop <- data.frame()
  anomalo_sn <- data.frame() 
  
  # Temporal aggregation (references period, d1 and d2, and temporal)
  # By cluster 
  for (i in variables){ 
    # summary for all pixels
    aux_sn <- mydf %>% 
      dplyr::group_by(month, event) %>%
      summarise_each_(funs(mean, sd, se=sd(.)/sqrt(n())), i) %>% mutate(variable=i)
    
    anomalo_sn <- rbind(anomalo_sn, aux_sn) 
    
    # summary for cluster pop   
    aux_clu <- mydf %>% 
      dplyr::group_by(clu_pop, month, event) %>%
      summarise_each_(funs(mean, sd, se=sd(.)/sqrt(n())), i) %>% mutate(variable=i) 
    
    anomalo_clu <- rbind(anomalo_clu, aux_clu)
    
    # summary for populations 
    aux_pop <- mydf %>% 
      dplyr::group_by(poblacion, month, event) %>%  
      summarise_each_(funs(mean, sd, se=sd(.)/sqrt(n())), i) %>% mutate(variable=i) 
    
    anomalo_pop <- rbind(anomalo_pop, aux_pop) 
    
    rm(aux_pop, aux_clu, aux_sn)
    }
  
  # Store df for each j of misdataframes
  assign(paste0("anomalo_sn", j), anomalo_sn)
  assign(paste0("anomalo_pop", j), anomalo_pop)
  assign(paste0("anomalo_clu", j), anomalo_clu)
  
  rm(anomalo_sn, anomalo_pop, anomalo_clu)
  rm(mydf)
}


# Exportdataframes 
write.csv(anomalo_snd, file=paste(di, "/data/anomalies/anomalo_sn.csv", sep=""), row.names = FALSE)
write.csv(anomalo_clud, file=paste(di, "/data/anomalies/anomalo_clu.csv", sep=""), row.names = FALSE)
write.csv(anomalo_popd, file=paste(di, "/data/anomalies/anomalo_pop.csv", sep=""), row.names = FALSE)

# Computar anomalias standarizadas 

conjuntodata <- c('anomalo_snd', 'anomalo_popd', 'anomalo_clud')
variables <- c('myevisp', 'myndvisp')
myyear <- c('2005', '2012')

for (g in conjuntodata){
  
  # Select conjunto
  midf <- get(g)
  
  # Loop by year 
  for (j in myyear) { 
    # filter by byear
    df <- midf %>%
      filter(event %in% c(j, 'ref')) 
    
    # empty df to store results 
    a <- data.frame() 
    
    # loop into variables
    for (i in variables){ 
      out <- df %>% 
        filter(variable == i) %>%
        select(-variable) %>% gather(variable, value, mean, se, sd) %>%
        unite(var, variable, event) %>% 
        spread(var, value) 
      
      out <- out %>% mutate(iv = i)
      
      a <- rbind(a, out) 
      }
    
    assign(paste0("a_",g,j), a)
  }
}



### Compute and Export anomalies as csv


#SN 
a_anomalo_snd2005 <- a_anomalo_snd2005 %>% 
  mutate(std_a = ((mean_2005 - mean_ref) / sd_ref)) 
write.csv(a_anomalo_snd2005, file=paste(di, "/data/anomalies/std_a_2005_sn.csv", sep=""), row.names = FALSE)

a_anomalo_snd2012 <- a_anomalo_snd2012 %>% 
  mutate(std_a = ((mean_2012 - mean_ref) / sd_ref)) 
write.csv(a_anomalo_snd2012, file=paste(di, "/data/anomalies/std_a_2012_sn.csv", sep=""), row.names = FALSE)


#Clu 
a_anomalo_clud2005 <- a_anomalo_clud2005 %>% 
  mutate(std_a = ((mean_2005 - mean_ref) / sd_ref)) 
write.csv(a_anomalo_clud2005, file=paste(di, "/data/anomalies/std_a_2005_clu.csv", sep=""), row.names = FALSE)

a_anomalo_clud2012 <- a_anomalo_clud2012 %>% 
  mutate(std_a = ((mean_2012 - mean_ref) / sd_ref)) 
write.csv(a_anomalo_clud2012, file=paste(di, "/data/anomalies/std_a_2012_clu.csv", sep=""), row.names = FALSE)



#Pop 
a_anomalo_popd2005 <- a_anomalo_popd2005 %>% 
  mutate(std_a = ((mean_2005 - mean_ref) / sd_ref)) 
write.csv(a_anomalo_popd2005, file=paste(di, "/data/anomalies/std_a_2005_pop.csv", sep=""), row.names = FALSE)

a_anomalo_popd2012 <- a_anomalo_popd2012 %>% 
  mutate(std_a = ((mean_2012 - mean_ref) / sd_ref)) 
write.csv(a_anomalo_popd2012, file=paste(di, "/data/anomalies/std_a_2012_pop.csv", sep=""), row.names = FALSE)
```










for (i in variables){ 
aux <- aux_d1 %>% 
  dplyr::group_by(poblacion, month, event) %>%  
  summarise_each_(funs(mean, sd, se=sd(.)/sqrt(n())), i) %>% mutate(variable=i) 

anomalo_pop <- rbind(anomalo_pop, aux) }

# For all population
variables <- c('myevisp', 'myndvisp')
anomalo_sn <- data.frame() 

for (i in variables){ 
aux <- aux_d1 %>% 
  dplyr::group_by(month, event) %>%  
  summarise_each_(funs(mean, sd, se=sd(.)/sqrt(n())), i) %>% mutate(variable=i) 

anomalo_sn <- rbind(anomalo_sn, aux) }
  

assign(paste0("anomalo_sn", j), anomalo_sn)
assign(paste0("anomalo_pop", j), anomalo_pop)
assign(paste0("anomalo_clu", j), anomalo_clu)
} 




  




```


## Annomalies for all *Q. pyrenaica* forests
```{r} 
label_variable <- c('myevisp' = 'EVI', 'myndvisp' = 'NDVI')  

label_pop <- c('1' = 'Lugros', '2' = 'Guejar-Sierra', '3' = 'Monachil',
                      '4' = 'Dilar', '5' = 'Durcal', '6' = 'Caniar', 
                      '7' = 'Poqueira', '8' = 'Trevelez')


## Plot for all populations 
anomalo_sn %>% 
  # dplyr::filter(variable == 'myevisp') %>% 
  ggplot(aes(x=as.factor(month), y=mean, group=event, colour=event)) + 
  geom_point() + geom_line() + theme_bw() + 
  xlab('Months') + ylab('') + 
  facet_wrap(~variable, 
             labeller = as_labeller(label_variable),
             scales = 'free_y') +
  theme(strip.background = element_rect(fill = "white"))
```


## Annomalies by populations
```{r, fig.height=14}
## Plot for each population 
# EVI
anomalo_pop %>% 
  dplyr::filter(variable == 'myevisp') %>% 
  ggplot(aes(x=as.factor(month), y=mean, group=event, colour=event)) + 
  geom_point() + geom_line() + theme_bw() + 
  xlab('Months') + ylab('EVI') + 
  facet_wrap(~poblacion, 
             labeller = as_labeller(label_pop)) +
  theme(strip.background = element_rect(fill = "white"))
```

```{r, fig.height=14}
# NDVI
anomalo_pop %>% 
  dplyr::filter(variable == 'myndvisp') %>% 
  ggplot(aes(x=as.factor(month), y=mean, group=event, colour=event)) + 
  geom_point() + geom_line() + theme_bw() + 
  xlab('Months') + ylab('NDVI') + 
  facet_wrap(~poblacion,
             labeller = as_labeller(label_pop)) +
  theme(strip.background = element_rect(fill = "white"))
```



## Annomalies by populations cluster
```{r, fig.height=14}
## Plot for each cluster
anomalo_clu %>% 
  ggplot(aes(x=as.factor(month), y=mean, group=event, colour=event)) + 
  geom_point() + geom_line() + theme_bw() + 
  xlab('Months') + ylab('') + 
  facet_grid(variable~clu_pop, 
             labeller = labeller(.rows=label_variable),
             scales = 'free_y') + 
  theme(strip.background = element_rect(fill = "white"))
```




