---
title: "Analysis resilience patterns two events"
author: "AJ Perez-Luque (@ajpelu)"
date: "2017 Jan"
output:  
  md_document:
    toc: true
    variant: markdown_github
---
```{r, echo=FALSE, message=FALSE}
require(knitr)
opts_chunk$set(fig.align='center', message = FALSE, echo = FALSE, warning = FALSE) 
```


# Exploring patterns of resilience components (models)

```{r}
#---------------------------------
# machine <- 'ajpelu'
machine <- 'ajpeluLap'
di <- paste('/Users/', machine, '/Dropbox/phd/phd_repos/qpyr_resilience', sep='')
#---------------------------------
```

```{r packages, warning=FALSE, message=FALSE}
source(paste0(di,"/R/exportpdf.R")) # function to export raster levelplots maps as pdf
library("dplyr")
library("ggplot2")
library("reshape2")
library("purrr")
library("pander")
library("knitr")
source("http://www.highstat.com/Book2/HighstatLibV6.R")
library("multcomp")
library("broom")
library("effects")
library("devtools")
# multicomp table 
devtools::source_gist("194e721077ec045a2b864088908e7aca", filename = "table_glht.r")
```


```{r}
# Read resilience
evi_resilience <- read.csv(file=paste(di, "/data/evi_resilience.csv", sep=""), header = TRUE, sep = ',')

# Prepare data
elev <- read.csv(file=paste(di, "/data/elev.csv", sep=""), header = TRUE, sep = ',')


eviresi <- evi_resilience %>%  
  dplyr::inner_join(elev, by='iv_malla_modi_id') %>% 
  mutate(clu_pop = as.factor(ifelse(pop %in% c(1,2,3,4,5), 'N', 'S'))) %>% 
  mutate(clu_popf = as.factor(ifelse(clu_pop == 'N', 'Northern slope', 'Southern slope'))) %>% 
  mutate(event = as.factor(ifelse(event == 1, 2005, 2012)))
  
```

# Obtain means values by cluster population
```{r}
variables <- c('dr','post', 'pre','rs','rc','rt','rrs')
myseasons <- unique(eviresi$seasonF)
auxdf <- data.frame()

for (m in myseasons){ 
  # filter by season
  aux_season <- eviresi %>% filter(seasonF == m)
  
  for (j in unique(eviresi$event)){
    # filter by event 
    auxilio <- aux_season %>% filter(event==j) 
    
    for (i in variables){ 
      aux <- auxilio %>% 
        dplyr::group_by(clu_pop) %>% 
        summarise_each_(funs(mean, sd, se=sd(.)/sqrt(n())), i) %>% 
        mutate(variable=i) %>% 
        mutate(event = j) %>% 
        mutate(seasonF = m)
      
      auxdf <- rbind(auxdf, aux)
    }
  }
}

auxdf <- auxdf %>% mutate(variable = plyr::mapvalues(variable,
                                                         c("dr","post","pre","rs","rc","rt","rrs"),
                                                         c("1_dr","2_post","0_pre","rs","rc","rt","rrs")))


label_variable <- c('rt' = 'Resistance', 
                    'rc' = 'Recovery',
                    'rs' = 'Resilience',
                    'rrs' = 'Relative Resilience',
                    '0_pre' = 'PreDrought',
                    '1_dr' = 'Drought',
                    '2_post' = 'PostDrought')
```



# Resilience 
## Mean values (Cluster population)
```{r}
mivariable <- 'rs'

# Model 
myformula <- as.formula(paste0(mivariable, " ~ event + clu_pop + event:clu_pop"))
mymodel <- aov(myformula, data=eviresi)
```

```{r}
auxdf %>% dplyr::filter(variable==mivariable) %>% pander(round=4, caption=paste0('Mean values (', mivariable,')'))
```

## Summary ANOVA model
```{r}
## Summary model 
tm <- broom::tidy(mymodel)

# See http://my.ilstu.edu/~wjschne/444/ANOVA.html#(1) 
pander(tm, round=5,caption = "ANOVA table", missing = '', 
       emphasize.strong.cells = 
               which(tm < 0.1 & tm == tm$p.value, arr.ind = T))
```             

```{r} 
gm <- broom::glance(mymodel)
gm <- apply(gm, 1, formatC, digits = 2, format = "f") %>% t()
colnames(gm) <- paste0("$",c("R^2","\\mathrm{adj}R^2","\\sigma_e","F","p","df_m","\\mathrm{logLik}","AIC","BIC","\\mathrm{dev}","df_e"),"$")
rownames(gm) <- "Statistic"
pander(t(gm))
```

## Effects plot 
### . ~ Cluster population 
```{r}
## Effects plots 
### Population 
plot(effect("clu_pop",mymodel), 
     main=paste0(mivariable, ' vs. Cluster population'),
     xlab='Cluster population')
```

### . ~ Disturbance Event
```{r}
### Elevation
plot(effect("event",mymodel), 
     main=paste0(mivariable, ' vs. Drought event'),
     xlab='Drought event')
```

### . ~ Cluster population:Elevation  
```{r} 
### Both 
plot(effect("event:clu_pop",mymodel), 
     main=paste0('Interaction plot (',mivariable, ')'),
     xlab='Drought event')
```

## Post-hoc comparison 
```{r}
## Multiple comparison 

tuk <- glht(mymodel, linfct = mcp(clu_pop = "Tukey"))

# Convert comparisons into letters 
df_letter <- fortify(cld(tuk)) %>%
  transmute(clu_pop = as.factor(lhs),
         tukey = letters) %>%
  mutate(variable = mivariable)

aux_name <- paste0('df_tuk_', mivariable)
assign(aux_name, df_letter)

# Get table 
mymult <- as.data.frame(table_glht(tuk))
pander(mymult, round=4,caption = "Post hoc comparison (Tukey, alpha = 0.05)", missing = '', 
       emphasize.strong.cells = 
               which(mymult < 0.05 & mymult == mymult[,4], arr.ind = T))
```


```{r}
## Multiple comparison 

tuk <- glht(mymodel, linfct = mcp(event = "Tukey"))

# Convert comparisons into letters 
df_letter <- fortify(cld(tuk)) %>%
  transmute(clu_pop = as.factor(lhs),
         tukey = letters) %>%
  mutate(variable = mivariable)

aux_name <- paste0('df_tuk_', mivariable)
assign(aux_name, df_letter)

# Get table 
mymult <- as.data.frame(table_glht(tuk))
pander(mymult, round=4,caption = "Post hoc comparison (Tukey, alpha = 0.05)", missing = '', 
       emphasize.strong.cells = 
               which(mymult < 0.05 & mymult == mymult[,4], arr.ind = T))
```

```{r}
library(lsmeans)
library(multcompView)

rr <- ref.grid(mymodel)

lsmeans(rr, "event")
lsmeans(rr, "clu_pop")
myls <- lsmeans(rr, "event", by ="clu_pop")
contrast(myls, method = "eff")
pairs(myls)

cld(myls)


summary(rr, infer = c(TRUE,TRUE), level = .90, adjust = "bon", by = "event")
lsmip(myls, event ~ clu_pop)

plot(myls, by='clu_pop')
# Resistence 

## Mean values (Cluster population)
```{r}
mivariable <- 'rt'

# Model 
myformula <- as.formula(paste0(mivariable, " ~ elev + clu_popf + elev:clu_popf"))
mymodel <- aov(myformula, data=eviresi_f)
```

```{r}
auxdf %>% dplyr::filter(variable==mivariable) %>% pander(round=4, caption=paste0('Mean values (', mivariable,')'))
```

## Summary ANCOVA model
```{r}
## Summary model 
tm <- broom::tidy(mymodel)

# See http://my.ilstu.edu/~wjschne/444/ANOVA.html#(1) 
pander(tm, round=5,caption = "ANOVA table", missing = '', 
       emphasize.strong.cells = 
         which(tm < 0.1 & tm == tm$p.value, arr.ind = T))
```             

```{r} 
gm <- broom::glance(mymodel)
gm <- apply(gm, 1, formatC, digits = 2, format = "f") %>% t()
colnames(gm) <- paste0("$",c("R^2","\\mathrm{adj}R^2","\\sigma_e","F","p","df_m","\\mathrm{logLik}","AIC","BIC","\\mathrm{dev}","df_e"),"$")
rownames(gm) <- "Statistic"
pander(t(gm))
```

## Effects plot 
### . ~ Cluster population 
```{r}
## Effects plots 
### Population 
plot(effect("clu_popf",mymodel), 
     main=paste0(mivariable, ' vs. Cluster population'),
     xlab='Cluster population')
```

### . ~ Elevation
```{r}
### Elevation
plot(effect("elev",mymodel), 
     main=paste0(mivariable, ' vs. Elevation'),
     xlab='Elevation (m)')
```

### . ~ Cluster population:Elevation  
```{r} 
### Both 
plot(effect("elev:clu_popf",mymodel), 
     main=paste0('Interaction plot (',mivariable, ')'),
     xlab='Elevation (m)')
```

## Post-hoc comparison 
```{r}
## Multiple comparison 

tuk <- glht(mymodel, linfct = mcp(clu_popf = "Tukey"))

# Convert comparisons into letters 
df_letter <- fortify(cld(tuk)) %>%
  transmute(clu_popf = as.factor(lhs),
            tukey = letters) %>%
  mutate(variable = mivariable)

aux_name <- paste0('df_tuk_', mivariable)
assign(aux_name, df_letter)

# Get table 
mymult <- as.data.frame(table_glht(tuk))
pander(mymult, round=4,caption = "Post hoc comparison (Tukey, alpha = 0.05)", missing = '', 
       emphasize.strong.cells = 
         which(mymult < 0.05 & mymult == mymult[,4], arr.ind = T))
```


# Recovery

## Mean values (Cluster population)
```{r}
mivariable <- 'rc'

# Model 
myformula <- as.formula(paste0(mivariable, " ~ elev + clu_popf + elev:clu_popf"))
mymodel <- aov(myformula, data=eviresi_f)
```

```{r}
auxdf %>% dplyr::filter(variable==mivariable) %>% pander(round=4, caption=paste0('Mean values (', mivariable,')'))
```

## Summary ANCOVA model
```{r}
## Summary model 
tm <- broom::tidy(mymodel)

# See http://my.ilstu.edu/~wjschne/444/ANOVA.html#(1) 
pander(tm, round=5,caption = "ANOVA table", missing = '', 
       emphasize.strong.cells = 
         which(tm < 0.1 & tm == tm$p.value, arr.ind = T))
```             

```{r} 
gm <- broom::glance(mymodel)
gm <- apply(gm, 1, formatC, digits = 2, format = "f") %>% t()
colnames(gm) <- paste0("$",c("R^2","\\mathrm{adj}R^2","\\sigma_e","F","p","df_m","\\mathrm{logLik}","AIC","BIC","\\mathrm{dev}","df_e"),"$")
rownames(gm) <- "Statistic"
pander(t(gm))
```

## Effects plot 
### . ~ Cluster population 
```{r}
## Effects plots 
### Population 
plot(effect("clu_popf",mymodel), 
     main=paste0(mivariable, ' vs. Cluster population'),
     xlab='Cluster population')
```

### . ~ Elevation
```{r}
### Elevation
plot(effect("elev",mymodel), 
     main=paste0(mivariable, ' vs. Elevation'),
     xlab='Elevation (m)')
```

### . ~ Cluster population:Elevation  
```{r} 
### Both 
plot(effect("elev:clu_popf",mymodel), 
     main=paste0('Interaction plot (',mivariable, ')'),
     xlab='Elevation (m)')
```

## Post-hoc comparison 
```{r}
## Multiple comparison 

tuk <- glht(mymodel, linfct = mcp(clu_popf = "Tukey"))

# Convert comparisons into letters 
df_letter <- fortify(cld(tuk)) %>%
  transmute(clu_popf = as.factor(lhs),
            tukey = letters) %>%
  mutate(variable = mivariable)

aux_name <- paste0('df_tuk_', mivariable)
assign(aux_name, df_letter)

# Get table 
mymult <- as.data.frame(table_glht(tuk))
pander(mymult, round=4,caption = "Post hoc comparison (Tukey, alpha = 0.05)", missing = '', 
       emphasize.strong.cells = 
         which(mymult < 0.05 & mymult == mymult[,4], arr.ind = T))
```

# Relative resilience

## Mean values (Cluster population)
```{r}
mivariable <- 'rrs'

# Model 
myformula <- as.formula(paste0(mivariable, " ~ elev + clu_popf + elev:clu_popf"))
mymodel <- aov(myformula, data=eviresi_f)
```

```{r}
auxdf %>% dplyr::filter(variable==mivariable) %>% pander(round=4, caption=paste0('Mean values (', mivariable,')'))
```

## Summary ANCOVA model
```{r}
## Summary model 
tm <- broom::tidy(mymodel)

# See http://my.ilstu.edu/~wjschne/444/ANOVA.html#(1) 
pander(tm, round=5,caption = "ANOVA table", missing = '', 
       emphasize.strong.cells = 
         which(tm < 0.1 & tm == tm$p.value, arr.ind = T))
```             

```{r} 
gm <- broom::glance(mymodel)
gm <- apply(gm, 1, formatC, digits = 2, format = "f") %>% t()
colnames(gm) <- paste0("$",c("R^2","\\mathrm{adj}R^2","\\sigma_e","F","p","df_m","\\mathrm{logLik}","AIC","BIC","\\mathrm{dev}","df_e"),"$")
rownames(gm) <- "Statistic"
pander(t(gm))
```

## Effects plot 
### . ~ Cluster population 
```{r}
## Effects plots 
### Population 
plot(effect("clu_popf",mymodel), 
     main=paste0(mivariable, ' vs. Cluster population'),
     xlab='Cluster population')
```

### . ~ Elevation
```{r}
### Elevation
plot(effect("elev",mymodel), 
     main=paste0(mivariable, ' vs. Elevation'),
     xlab='Elevation (m)')
```

### . ~ Cluster population:Elevation  
```{r} 
### Both 
plot(effect("elev:clu_popf",mymodel), 
     main=paste0('Interaction plot (',mivariable, ')'),
     xlab='Elevation (m)')
```

## Post-hoc comparison 
```{r}
## Multiple comparison 

tuk <- glht(mymodel, linfct = mcp(clu_popf = "Tukey"))

# Convert comparisons into letters 
df_letter <- fortify(cld(tuk)) %>%
  transmute(clu_popf = as.factor(lhs),
            tukey = letters) %>%
  mutate(variable = mivariable)

aux_name <- paste0('df_tuk_', mivariable)
assign(aux_name, df_letter)

# Get table 
mymult <- as.data.frame(table_glht(tuk))
pander(mymult, round=4,caption = "Post hoc comparison (Tukey, alpha = 0.05)", missing = '', 
       emphasize.strong.cells = 
         which(mymult < 0.05 & mymult == mymult[,4], arr.ind = T))
```

# Summary plots 

```{r, fig.width=14}
df_all_letters <- rbind(df_tuk_rs, df_tuk_rt, df_tuk_rc, df_tuk_rrs)

auxdf <- auxdf %>% 
  inner_join(df_all_letters, by=c('clu_popf', 'variable')) 


gpop_bar_letter <- ggplot(auxdf[auxdf$variable != 'rrs',], aes(x=clu_popf, y=mean)) + 
  geom_bar(stat='identity', fill='black', colour='black') + 
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), 
                width=.2, position=position_dodge(.9))  +
  facet_wrap(~variable, labeller = as_labeller(label_variable)) + 
  theme_bw() + xlab('') + ylab('') + 
  geom_text(aes(label = tukey, 
                y=(mean+1.7*sd)),
            position = position_dodge(width=0.9)) +
  theme(strip.background = element_rect(fill = "white"))

gpop_bar_letter
pdf(file=paste0(di, "/man/images/plot_resicomp_bar_tukey_by_cluster.pdf"), height = 7, width = 8)
gpop_bar_letter
dev.off()
```

```{r, fig.width=6}
gpop_bar_letter_rrs <- ggplot(auxdf[auxdf$variable == 'rrs',], aes(x=clu_popf, y=mean)) + 
  geom_bar(stat='identity', fill='black', colour='black') + 
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), 
                width=.2, position=position_dodge(.9))  +
  theme_bw() + xlab('') + ylab('Relative resilience') + 
  geom_text(aes(label = tukey, 
                y=(mean+1.7*sd)),
            position = position_dodge(width=0.9)) +
  theme(strip.background = element_rect(fill = "white"))

gpop_bar_letter_rrs
pdf(file=paste0(di, "/man/images/plot_rrs_bar_tukey_by_cluster.pdf"), height = 7, width = 8)
gpop_bar_letter_rrs
dev.off()
```


