---
title: "Resilience of Q. pyrenaica to a drought event"
author: "AJ Perez-Luque (@ajpelu)"
date: "2016 May"
output:  
  md_document:
    variant: markdown_github
---
# Compute resilience values of *Q. pyrenaica* for EVI values 

```{r wd, echo=FALSE}
#---------------------------------
machine <- 'ajpelu'
# machine <- 'ajpeluLap'
di <- paste('/Users/', machine, '/Dropbox/phd/phd_repos/qpyr_resilience', sep='')
#---------------------------------
```

```{r packages, warning=FALSE, message=FALSE}
library("dplyr")
library("ggplot2")
library("reshape2")
```

## Introduction 

* Read data of EVI attributes (see this [script](/analysis/prepare_evi_data.md)) 

```{r}
# Read data
evidf <- read.csv(file=paste(di, "/data/evi_attributes_all.csv", sep=""), header = TRUE, sep = ',')
```

## Exploring analysis evolution of EVI by season

* First we explore the evolution of EVI by season. We want to ask the question: *¿Which temporal scale does capture better the impacts of the drought event?*  

```{r, fig.width=12, fig.height=8, fig.cap="Evolution of seasonal and annual EVI for Q. pyrenaica forests in S.Nevada (2000-2014)"}
ggplot(evidf, aes(x=year, y=evi, group=year)) + 
  geom_boxplot() + 
  facet_wrap(~seasonF, scales = "free_y") + 
  theme_bw() + 
  theme(strip.background=element_rect(fill='white')) + 
  ylab('EVI')


evidf %>% 
  filter(seasonF %in% c('spring','summer')) %>% 
  group_by(year) %>% 
  ggplot(aes(x=year, y=evi, group=year)) +
  geom_boxplot() + theme_bw() + 
  facet_wrap(~seasonF) +
  theme(strip.background=element_rect(fill='white')) + 
  ylab('EVI')
```

## Define the drought period 

* Another key step is to define the number of drought events and the **time span** of the events. For this purpose we review some literature ([^Garcia_Herrera], [^Trigo_et_al_2013], [^Gouveia_and_Trigo]). Some notes:  

> Note: Drougths events.
>
> According to García-Herrera et al. 2007, the 6 greater droughts events recorded in Granada station from 1941 to 2005 were: 44-45, 48-49, 49-50, 92-93, 94-95, and 98-99. 
>
> Trigo et al. 2013 reported a winter drought event for souther Iberian in 2011-2012. 
>
> During the hydrological years of 2004/2005 and 2011/2012, Iberia was hit by two of the worst drought episodes ever recording in this semi-arid region (Gouveia and Trigo 2014)

We filter 2000-2002 and > 2010, to revmoe the potential bias produced by this drougt events. 

We states the period as: 

* Drought event: `2005-2006`
* preDrought: `2003-2004`
* postDrought: `2007-2010`


```{r}
# define period 
pre <- c(2003,2004)
dr <- c(2005,2006)
post <- c(2007,2008,2009,2010)
```

Assign, also, names for each *Q. pyrenaica* patch:  

```{r}
# names of populations
label_pop <- c('1' = 'Lugros', '2' = 'Guejar-Sierra', '3' = 'Monachil',
                      '4' = 'Dilar', '5' = 'Durcal', '6' = 'Caniar', 
                      '7' = 'Poqueira', '8' = 'Trevelez', '9' = 'Cadiar')
```

## Explore evolution of summer EVI

* Explore evolution of EVI summer 
* Filter data to period consider (`min(pre)` - `max(post)`)
* Plot the evolution of EVI summer 
* Categorize summer EVI value into `pre`, `dr` and `post` leveles according to drought events defined above

```{r, fig.width=12, fig.height=10, fig.cap="Evolution of summer EVI for Q. pyrenaica forests in S.Nevada, from 2003 to 2010"}
# Choose season
myseason <- 'summer'

evisummer <- evidf %>%
  filter(seasonF == myseason) %>% 
  mutate(myperiod = ifelse(year %in% pre, 'pre',
                         ifelse(year %in% dr, 'dr',
                           ifelse(year %in% post, 'post', 'out')))) %>%
  filter(myperiod != 'out') 



ggplot(evisummer, aes(x=year, y=evi, group=year)) + 
  geom_boxplot() + facet_wrap(~poblacion, labeller = as_labeller(label_pop)) + 
  ggtitle(paste0(myseason, ' EVI')) + 
  theme_bw() + 
  theme(strip.background=element_rect(fill='white'))
```

## Compute resilience components 

* Compute the resiliencie components according to (Lloret et al. 2011)([^Lloret_et_al]) (see also this [notes](/man/notes_drought_in.md)): ***Resistence*** (`rt`); ***Resilience*** (`rs`) and ***Recovery*** (`rc`). 
* These inidcators were computed at pixel scale. For this purpose we aggregated the values of each period by pixel (*i.e.*: mean values of summer EVI for several years) 
* We also retain the attributes of `population` and coordinates (`lat` and `lng`) 
* Finally we export the dataframe generate ([`/data/evi_resilience.csv`](/data/evi_resilience.csv))

```{r}
# First aggregate each pixel by period (get mean value of evi by period) 
eviperiod <- evisummer %>% 
  select(iv_malla_modi_id, poblacion, evi, myperiod, lng, lat) %>%
  group_by(iv_malla_modi_id, myperiod, poblacion, lng, lat) %>% 
  summarise(media_period = mean(evi))

# Change format of the dataset (long to wide)
df_aux <- dcast(eviperiod, iv_malla_modi_id + poblacion + lng + lat ~ myperiod, value.var = 'media_period')
  
# Compute components of resilience 

eviresi <- df_aux %>% 
  mutate(rt = dr / pre,
         rc = post / dr,
         rs = post / pre) %>%
  mutate(rrs = ((post - dr) / pre))

# Export data
write.csv(eviresi, file=paste(di, "/data/evi_resilience.csv", sep=""), row.names = FALSE)
```


## References

* [^Garcia_Herrera]: Garcia-Herrera R, Paredes D, Trigo RM, Trigo IF, Hernandez E, Barriopedro D, Mendes MA (2007). The Outstanding 2004/05 Drought in the Iberian Peninsula: Associated Atmospheric Circulation. *Journal of Hydrometeorology*, 8: 483–498.

* [^Gouveia_and_Trigo]: Gouveia CM and Trigo RM (2014) The 2005 and 2012 major drought events in Iberia: monitoring vegetation dynamics and crop yields using satellite data. EGU General Assembly Conference. Vol 16, EGU2014-15179-1. http://meetingorganizer.copernicus.org/EGU2014/EGU2014-15179-1.pdf 

* [^Trigo_et_al_2013]: Trigo RM, Añel J, Barriopedro D, García-Herrera R, Gimeno L, Nieto R, Castillo R, Allen MR, Massey N (2013). The record Winter drought of 2011-12 in the Iberian Peninsula [in *"Explaining Extreme Events of 2012 from a Climate Perspective"*. [Peterson TC, Hoerling MP, Stott PA Herring S (Eds.)]. *Bulletin of the American Meteorological Society*, 94 (9): S41-S45.

* [^Lloret_et_al]: Lloret F, Keeling EG, Sala A (2011) Components of tree resilience: effects of successive low-growth episodes in old ponderosa pine forests. *Oikos* 120: 1909–1920. 



```{r, eval=FALSE, echo=FALSE}

ggplot(eviresi, aes(x=as.factor(poblacion), y=rc, group=poblacion)) + 
  geom_boxplot() 
  
  
  facet_wrap(~poblacion, labeller = as_labeller(label_pop)) +
  theme_bw() + 
  xlab('') + ylab('summer EVI (average)') + 
  theme(strip.background=element_rect(fill='white'))





# Aggregate by period 
df_aux <- df %>% 
  group_by(iv_malla_modi_id, myperiod, poblacion) %>% 
  summarise(media = mean(evi)) %>% 
  mutate(myperiod_num = ifelse(myperiod == 'pre', 0,
                               ifelse(myperiod == 'dr', 1, 2)))

df_aux$myperiod_o <- factor(df_aux$myperiod, c('pre', 'dr', 'post'))

ggplot(df_aux, aes(x=myperiod_o, y=media, group=myperiod)) + 
  geom_boxplot() + facet_wrap(~poblacion, labeller = as_labeller(label_pop)) +
  theme_bw() + 
  xlab('') + ylab('summer EVI (average)') + 
  theme(strip.background=element_rect(fill='white'))







# Create mean values by pob and year
mydf <- evi %>% 
  filter(seasonF == 'summer') %>%
  select(year, poblacion, evi_mean) %>%
  group_by(year, poblacion) %>% 
  summarise(media = mean(evi_mean), 
            sd = sd(evi_mean), 
            se = sd(evi_mean) / sqrt(length(evi_mean))) 
                                         
  

ggplot(mydf, aes(x=year, y=media, group=year)) + 
  geom_point() +
  geom_errorbar(aes(ymin=media - se, ymax= media + se)) +
  facet_wrap(~poblacion) + 
  geom_line(aes(group=1))






# Resistance 
# rt = dr / pre 
# Recovery 
# rc = post / dr 
# Resilience 
# rs = post /pre 


df <- evi %>%
  mutate(myperiod = ifelse(year %in% c(2003,2004), 
                         'pre', 
                         ifelse( 
                           year %in% c(2005), 
                           'dr',
                           ifelse(
                             year %in% c(2006,2007), 
                             'post', 'out'))))

df <- df %>% filter(myperiod != 'out') %>%
  select(iv_malla_modi_id, poblacion, evi_mean_all, myperiod) %>%
  group_by(iv_malla_modi_id, myperiod) %>% 
  summarise(media = mean(evi_mean_all))


# long to wide 
library(reshape2)
df_aux <- dcast(df, iv_malla_modi_id ~ myperiod, value.var = 'media')

# Resistance rt = dr / pre 
# Recoveryrc = post / dr 
# Resilience rs = post /pre 



df_aux <- df_aux %>% 
  mutate(rt = dr / pre,
         rc = post / dr,
         rs = post / pre) %>% 
  inner_join(evi_aux, by="iv_malla_modi_id") 


ggplot(df_aux, aes(x=year, y=evi_mean, group=year)) + geom_boxplot() + facet_wrap(~poblacion)


```







