---
title: "Analysis annomalies"
author: "AJ Perez-Luque (@ajpelu)"
date: "2016 Aug"
output:  
  md_document:
    variant: markdown_github
---
```{r, echo=FALSE, message=FALSE}
require(knitr)
opts_chunk$set(fig.align='center', fig.width = 14, message = FALSE, warning = FALSE) 
```

```{r wd, echo=FALSE}
#---------------------------------
#machine <- 'ajpelu'
machine <- 'ajpeluLap'
di <- paste('/Users/', machine, '/Dropbox/phd/phd_repos/qpyr_resilience', sep='')
#---------------------------------
```


```{r packages, warning=FALSE, message=FALSE}
library("dplyr")
library("lubridate")
library("ggplot2")
library("RCurl")
library("tidyr")
```

# Read data 
Two dataframes: 
* EVI seasonal 
* IV composites 

```{r}
# Read data
iv <- read.csv(file=paste(di, "/data/iv_composite.csv", sep=""), header = TRUE, sep = ',')
evi <- read.csv(file=paste(di, "/data/evi_atributes_all.csv", sep=""), header = TRUE, sep = ',')
```


# Compute anomalies 
## By composite

```{r, eval=FALSE}
# Set counters
pixels <- unique(iv$iv_malla_modi_id)
composites <- unique(iv$composite)
years <- unique(iv$year)

# Set dataframe to compute anomaly
df <- iv 

# Create emtpy datafraje to store output 
anomalos <- data.frame()

for (i in pixels){
  df_aux <- df[df$iv_malla_modi_id == i,]
  
  # Create empty auxiliar dataframe to store results
  aux_composite <- data.frame() 
  
  for (j in composites) {
    # Create df_auxiliar by composite
    df_by_composite <- df_aux[df_aux$composite == j,]
    
    # Create empty df to store anomlay of composite j of all year 
    aux_anomaly <- data.frame() 
    
    for (y in years){
      # get mean evi for the reference period
      iv_ref <- df_by_composite %>% 
        filter(year != y) %>% 
        summarise(mean(evi))
      
      # get evi for the year
      iv_year <- df_by_composite[df_by_composite$year == y, 'evi']
      # To solve the problems with the year without composites (i.e 2000)
      iv_year <- ifelse(length(iv_year) == 0, 0, iv_year)
      
      # Compute standardized anomaly
      anomaly_std <- ((iv_year - iv_ref) / (iv_year + iv_ref))*100
      names(anomaly_std) <- 'anomaly_std'
      
      # Compute anomaly
      anomaly <- (iv_year - iv_ref)
      names(anomaly) <- 'anomaly'
      
      # Create dataframe 
      aux <- cbind(y, anomaly, anomaly_std)
      aux_anomaly <- rbind(aux_anomaly, aux)
      } 
      
    aux_compos <- aux_anomaly %>% mutate(composite = j)
    
    aux_composite <- rbind(aux_composite, aux_compos)
    
  }
    
  # Remove the composites 1,2 an 3 for year 2000 
  aux_composite <- aux_composite[!(aux_composite$y == 2000 & aux_composite$composite %in% c(1:3)), ] 
    
  # Add name of the pixel 
  aux_composite <- aux_composite %>% mutate(iv_malla_modi_id = i)
  
  # Output 
  anomalos <- rbind(anomalos, aux_composite)
}   


# Add pob, lat, long, etc 
iv_aux <- iv %>% dplyr::select(iv_malla_modi_id, long, lat, pop) %>%
  group_by(iv_malla_modi_id) %>% unique()

# Join dataframes 
anomalias_composite <- anomalos %>% dplyr::inner_join(iv_aux, by="iv_malla_modi_id") 

write.csv(anomalias_composite, file=paste(di, "/data/anomalies/anomalias_composite.csv", sep=""), row.names = FALSE)
```

# By season

```{r, eval=FALSE}
# Set counters
pixels <- unique(evi$iv_malla_modi_id)
composites <- unique(evi$seasonF)
years <- unique(evi$year)

# Set dataframe to compute anomaly
df <- evi

# Create emtpy datafraje to store output 
anomalos <- data.frame()

for (i in pixels){
  df_aux <- df[df$iv_malla_modi_id == i,]
  
  # Create empty auxiliar dataframe to store results
  aux_composite <- data.frame() 
  
  for (j in composites) {
    # Create df_auxiliar by composite
    df_by_composite <- df_aux[df_aux$seasonF == j,]
    
    # Create empty df to store anomlay of composite j of all year 
    aux_anomaly <- data.frame() 
    
    for (y in years){
      # get mean evi for the reference period
      iv_ref <- df_by_composite %>% 
        filter(year != y) %>% 
        summarise(mean(evi))
      
      # get evi for the year
      iv_year <- df_by_composite[df_by_composite$year == y, 'evi']
      # To solve the problems with the year without composites (i.e 2000)
      iv_year <- ifelse(length(iv_year) == 0, 0, iv_year)
      
      # Compute standardized anomaly
      anomaly_std <- ((iv_year - iv_ref) / (iv_year + iv_ref))*100
      names(anomaly_std) <- 'anomaly_std'
      
      # Compute anomaly
      anomaly <- (iv_year - iv_ref)
      names(anomaly) <- 'anomaly'
      
      # Create dataframe 
      aux <- cbind(y, anomaly, anomaly_std)
      aux_anomaly <- rbind(aux_anomaly, aux)
      } 
      
    aux_compos <- aux_anomaly %>% mutate(composite = j)
    
    aux_composite <- rbind(aux_composite, aux_compos)
    
  }
    
  # Add name of the pixel 
  aux_composite <- aux_composite %>% mutate(iv_malla_modi_id = i)
  
  # Output 
  anomalos <- rbind(anomalos, aux_composite)
}   



# Add pob, lat, long, etc 
evi_aux <- evi %>% dplyr::select(iv_malla_modi_id, long, lat, pop) %>%
  group_by(iv_malla_modi_id) %>% unique()

# Join dataframes 
anomalias_season <- anomalos %>% dplyr::inner_join(evi_aux, by="iv_malla_modi_id") 

write.csv(anomalias_season, file=paste(di, "/data/anomalies/anomalias_season.csv", sep=""), row.names = FALSE)
```




        

          
      

        

        


  





# Set disturbance years 
disturb_years <- c(2005, 2012)








d <- rawdata %>% 
    # Factor Correction to evi and ndvi
    mutate(myevi = evi * 0.0001,
           myndvi = ndvi * 0.0001) %>% 
    select(-evi) %>% 
    select(-ndvi) %>% 
    
    # Compute year and month 
    mutate(year = lubridate::year(date),  
           month = lubridate::month(date),
           yearday = lubridate::yday(date)) %>%

    
    # Add variable of population cluster and remove population '9'
    mutate(clu_pop = as.factor(ifelse(poblacion == 1, 'Camarate',
                                      ifelse(poblacion %in% c(2,3,4,5), 'Northern slope',
                                             ifelse(poblacion %in% c(6,7,8), 'Southern slope', 'out'))))) %>% 
    filter(clu_pop != 'out') %>% 
  
      # Categorize reference period and disturbance year
    mutate(event = ifelse(year != disturb_years,
                          'ref',
                          ifelse(year == disturb_years[1], 
                                 as.character(disturb_years[1]),
                                 as.character(disturb_years[2])))) %>% 
  
      # Spatial aggregation. All pixels of the same populations are aggregate by mean
    group_by(iv_malla_modi_id, clu_pop, poblacion, year, month, yearday, event) %>% 
    summarise(myevisp = mean(myevi), 
              myndvisp = mean(myndvi)) 
  
















d <- rawdata %>% 
    # Factor Correction to evi and ndvi
    mutate(myevi = evi * 0.0001,
           myndvi = evi * 0.0001) %>% 
    select(-evi) %>% 
    select(-ndvi) %>% 
    
    # Compute year and month 
    mutate(year = lubridate::year(date),  
           month = lubridate::month(date)) %>% 
    
    # Add variable of population cluster and remove population '9'
    mutate(clu_pop = as.factor(ifelse(poblacion == 1, 'Camarate',
                                      ifelse(poblacion %in% c(2,3,4,5), 'Northern slope',
                                             ifelse(poblacion %in% c(6,7,8), 'Southern slope', 'out'))))) %>% 
    filter(clu_pop != 'out') %>% 
    
    # Categorize reference period and disturbance year
    mutate(event = ifelse(year != disturb_years,
                          'ref',
                          ifelse(year == disturb_years[1], 
                                 as.character(disturb_years[1]),
                                 as.character(disturb_years[2])))) %>% 
    
    # Spatial aggregation. All pixels of the same populations are aggregate by mean
    group_by(clu_pop, poblacion, year, month, event) %>% 
    summarise(myevisp = mean(myevi), 
              myndvisp = mean(myndvi)) 
  }
```



```{r}

# Note 

computeAnomaly <- function(df, disturb_years){
  require('dplyr')
  
  
  rd <- df %>% 
    # Factor Correction to evi and ndvi
    mutate(myevi = evi * 0.0001,
           myndvi = evi * 0.0001) %>% 
    select(-evi) %>% 
    select(-ndvi) %>% 
    
    # Compute year and month 
    mutate(year = lubridate::year(date),  
           month = lubridate::month(date)) %>% 
    
    # Add variable of population cluster and remove population '9'
    mutate(clu_pop = as.factor(ifelse(poblacion == 1, 'Camarate',
                                      ifelse(poblacion %in% c(2,3,4,5), 'Northern slope',
                                             ifelse(poblacion %in% c(6,7,8), 'Southern slope', 'out'))))) %>% 
    filter(clu_pop != 'out') %>% 
    
    # Categorize reference period and disturbance year
    mutate(event = ifelse(year != disturb_years,
                          'ref',
                          ifelse(year == disturb_years[1], 
                                 as.character(disturb_years[1]),
                                 as.character(disturb_years[2])))) %>% 
    
    # Spatial aggregation. All pixels of the same populations are aggregate by mean
    group_by(clu_pop, poblacion, year, month, event) %>% 
    summarise(myevisp = mean(myevi), 
              myndvisp = mean(myndvi)) 
  }


d <- computeAnomaly(df=rawdata, disturb_years = c(2005, 2012))
# d2 <- computeAnomaly(df=rawdata, disturb_year = 2012)
  


### Loop for temporal aggregation 

# misdataframes <- c("d1", "d2")
misdataframes <- c("d")
variables <- c('myevisp', 'myndvisp')

for (j in misdataframes){ 
  
  mydf <- get(j)
  
  # Create three empty dataframe
  anomalo_clu <- data.frame() 
  anomalo_pop <- data.frame()
  anomalo_sn <- data.frame() 
  
  # Temporal aggregation (references period, d1 and d2, and temporal)
  # By cluster 
  for (i in variables){ 
    
    #all_var_interest <- c('clu_pop', 'year', 'month', 'event', i)  
    
    # mydf <- mydf %>% dplyr::select(one_of(dplyr::intersect(all_var_interest, names(.))))

    
    # summary for all pixels
    aux_sn <- mydf %>% 
      dplyr::group_by(month, event) %>%
      dplyr::summarise_each_(funs(mean, sd, se=sd(.)/sqrt(n())), i) %>% mutate(variable=i)
    
    anomalo_sn <- rbind(anomalo_sn, aux_sn) 
    
    # summary for cluster pop   
    aux_clu <- mydf %>% 
      dplyr::group_by(clu_pop, month, event) %>%
      dplyr::summarise_each_(funs(mean, sd, se=sd(.)/sqrt(n())), i) %>% mutate(variable=i) 
    
    anomalo_clu <- rbind(anomalo_clu, aux_clu)
    
    # summary for populations 
    aux_pop <- mydf %>% 
      dplyr::group_by(poblacion, month, event) %>%  
      dplyr::summarise_each_(funs(mean, sd, se=sd(.)/sqrt(n())), i) %>% mutate(variable=i) 
    
    anomalo_pop <- rbind(anomalo_pop, aux_pop) 
    
    rm(aux_pop, aux_clu, aux_sn)
    }
  
  # Store df for each j of misdataframes
  assign(paste0("anomalo_sn", j), anomalo_sn)
  assign(paste0("anomalo_pop", j), anomalo_pop)
  assign(paste0("anomalo_clu", j), anomalo_clu)
  
  rm(anomalo_sn, anomalo_pop, anomalo_clu)
}


# Exportdataframes 
write.csv(anomalo_snd, file=paste(di, "/data/anomalies/anomalo_sn.csv", sep=""), row.names = FALSE)
write.csv(anomalo_clud, file=paste(di, "/data/anomalies/anomalo_clu.csv", sep=""), row.names = FALSE)
write.csv(anomalo_popd, file=paste(di, "/data/anomalies/anomalo_pop.csv", sep=""), row.names = FALSE)

# Computar anomalias standarizadas 

conjuntodata <- c('anomalo_snd', 'anomalo_popd', 'anomalo_clud')
variables <- c('myevisp', 'myndvisp')
myyear <- c('2005', '2012')

for (g in conjuntodata){
  
  # Select conjunto
  midf <- get(g)
  
  # Loop by year 
  for (j in myyear) { 
    # filter by byear
    df <- midf %>%
      filter(event %in% c(j, 'ref')) 
    
    # empty df to store results 
    a <- data.frame() 
    
    # loop into variables
    for (i in variables){ 
      out <- df %>% 
        filter(variable == i) %>%
        select(-variable) %>% gather(variable, value, mean, se, sd) %>%
        unite(var, variable, event) %>% 
        spread(var, value) 
      
      out <- out %>% mutate(iv = i)
      
      a <- rbind(a, out) 
      }
    
    assign(paste0("a_",g,j), a)
  }
}



### Compute and Export anomalies as csv


#SN 
a_anomalo_snd2005 <- a_anomalo_snd2005 %>% 
  mutate(std_a = ((mean_2005 - mean_ref) / sd_ref)) 
write.csv(a_anomalo_snd2005, file=paste(di, "/data/anomalies/std_a_2005_sn.csv", sep=""), row.names = FALSE)

a_anomalo_snd2012 <- a_anomalo_snd2012 %>% 
  mutate(std_a = ((mean_2012 - mean_ref) / sd_ref)) 
write.csv(a_anomalo_snd2012, file=paste(di, "/data/anomalies/std_a_2012_sn.csv", sep=""), row.names = FALSE)


#Clu 
a_anomalo_clud2005 <- a_anomalo_clud2005 %>% 
  mutate(std_a = ((mean_2005 - mean_ref) / sd_ref)) 
write.csv(a_anomalo_clud2005, file=paste(di, "/data/anomalies/std_a_2005_clu.csv", sep=""), row.names = FALSE)

a_anomalo_clud2012 <- a_anomalo_clud2012 %>% 
  mutate(std_a = ((mean_2012 - mean_ref) / sd_ref)) 
write.csv(a_anomalo_clud2012, file=paste(di, "/data/anomalies/std_a_2012_clu.csv", sep=""), row.names = FALSE)



#Pop 
a_anomalo_popd2005 <- a_anomalo_popd2005 %>% 
  mutate(std_a = ((mean_2005 - mean_ref) / sd_ref)) 
write.csv(a_anomalo_popd2005, file=paste(di, "/data/anomalies/std_a_2005_pop.csv", sep=""), row.names = FALSE)

a_anomalo_popd2012 <- a_anomalo_popd2012 %>% 
  mutate(std_a = ((mean_2012 - mean_ref) / sd_ref)) 
write.csv(a_anomalo_popd2012, file=paste(di, "/data/anomalies/std_a_2012_pop.csv", sep=""), row.names = FALSE)
```

