---
title: "Explore resilience patterns"
author: "AJ Perez-Luque (@ajpelu)"
date: "2016 May"
output:  
  md_document:
    variant: markdown_github
---
# Explore resilience patterns of the different *Q. pyrenaica* patches

```{r wd, echo=FALSE}
#---------------------------------
#machine <- 'ajpelu'
machine <- 'ajpeluLap'
di <- paste('/Users/', machine, '/Dropbox/phd/phd_repos/qpyr_resilience', sep='')
#---------------------------------
```

```{r packages, warning=FALSE, message=FALSE}
library("rgdal")
library("sp")
library("raster")
library("rasterVis")
source(paste0(di,"/R/exportpdf.R")) # function to export raster levelplots maps as pdf
library("dplyr")
library("ggplot2")
library("reshape2")
library("purrr")
# library("pander")
library("knitr")
```

## Read and prepare data 

```{r}
# Read Summer resilience
evi_resilience <- read.csv(file=paste(di, "/data/evi_resilience.csv", sep=""), header = TRUE, sep = ',')
```


# Maps 

* Create raster maps (two projections: `epsg:4326` and `epsg:23030`)
* Export raster maps and stacks (see `/data/raster/`)

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Create a spatial dataframe 
df_sel <- SpatialPointsDataFrame(coords = evi_resilience[,c('long','lat')], data = evi_resilience,
                               proj4string = CRS("+init=epsg:4326"))

# Transform
df_sel_sp <- spTransform(df_sel, CRS("+init=epsg:23030"))
  
myevent <- unique(evi_resilience$event)
seasons <- unique(evi_resilience$seasonF)
resilience_indicators <- c('rs','rt', 'rc', 'rrs')

for (j in myevent){
  # Filter my event 
  aux_spatial_event <- df_sel_sp[df_sel_sp$event == j,] # aux_spatial_year = aux_spatial_event
  
  for (m in seasons){
    # Filter my season 
    aux_spatial_event_season <- aux_spatial_event[aux_spatial_event$seasonF == m,]
    
    for (i in resilience_indicators) { 
      aux_spatial <- aux_spatial_event_season 
      
      # raster auxiliar 
      aux_rast <- raster(aux_spatial, resolution=250)
      # variable z
      zvar <- i
      
      # raster_layer 
      r_resi <- rasterize(aux_spatial, aux_rast, i, fun=mean)
      names(r_resi) <- paste0(i, "_", m, "_", j, sep="")
      
      # reprojected raster
      r_resi_re <- projectRaster(r_resi, crs=crs("+init=epsg:4326"))
      names(r_resi_re) <- paste0(i, "_", m, "_", j, sep="")
      
      # assign 
      name_raster <- paste("r_", i, "_", m, "_", j, sep="")
      name_raster_re <- paste("r_",i,"_re", "_", m, "_", j, sep="")
      
      assign(name_raster, r_resi)
      assign(name_raster_re, r_resi_re)
      
      writeRaster(r_resi, file=paste(di, "/data/raster/r_", i, "_", m, "_", j, ".asc", sep=""), overwrite=TRUE)
      writeRaster(r_resi_re, file=paste(di, "/data/raster/r_", i, "_re", "_", m, "_", j, ".tif", sep=""), overwrite=TRUE)
      
      
    }
  }
}


# Create stacks
stack_rc <- stack(r_rc_annual_1, r_rc_annual_2, r_rc_spring_1, r_rc_spring_2, r_rc_summer_1, r_rc_summer_2)
stack_rc_re <- stack(r_rc_re_annual_1, r_rc_re_annual_2, r_rc_re_spring_1, r_rc_re_spring_2, r_rc_re_summer_1, r_rc_re_summer_2)

stack_rs <- stack(r_rs_annual_1, r_rs_annual_2, r_rs_spring_1, r_rs_spring_2, r_rs_summer_1, r_rs_summer_2)
stack_rs_re <- stack(r_rs_re_annual_1, r_rs_re_annual_2, r_rs_re_spring_1, r_rs_re_spring_2, r_rs_re_summer_1, r_rs_re_summer_2)

stack_rt <- stack(r_rt_annual_1, r_rt_annual_2, r_rt_spring_1, r_rt_spring_2, r_rt_summer_1, r_rt_summer_2)
stack_rt_re <- stack(r_rt_re_annual_1, r_rt_re_annual_2, r_rt_re_spring_1, r_rt_re_spring_2, r_rt_re_summer_1, r_rt_re_summer_2)

stack_rrs <- stack(r_rrs_annual_1, r_rrs_annual_2, r_rrs_spring_1, r_rrs_spring_2, r_rrs_summer_1, r_rrs_summer_2)
stack_rrs_re <- stack(r_rrs_re_annual_1, r_rrs_re_annual_2, r_rrs_re_spring_1, r_rrs_re_spring_2, r_rrs_re_summer_1, r_rrs_re_summer_2)


# Export stacks 
temp <- getwd()
setwd(paste(di, "/data/raster/", sep=""))
writeRaster(stack_rc, filename = 'r_stack_rc', overwrite =TRUE) 
writeRaster(stack_rc_re, filename = 'r_stack_rc_re', overwrite =TRUE) 
writeRaster(stack_rs, filename = 'r_stack_rs', overwrite =TRUE) 
writeRaster(stack_rs_re, filename = 'r_stack_rs_re', overwrite =TRUE) 
writeRaster(stack_rt, filename = 'r_stack_rt', overwrite =TRUE) 
writeRaster(stack_rt_re, filename = 'r_stack_rt_re', overwrite =TRUE) 
writeRaster(stack_rrs, filename = 'r_stack_rrs', overwrite =TRUE) 
writeRaster(stack_rrs_re, filename = 'r_stack_rrs_re', overwrite =TRUE) 
setwd(temp)
```  
    
    

## Spatial exploration of the resilience components 

### Resilience
```{r, echo=FALSE, fig.cap='Resilience', fig.height=7, fig.width=10}
# Resilience 

# Select a palette http://colorbrewer2.org/
mypal <- brewer.pal(9, "RdYlGn")
# Specify the color palette
myTheme=rasterTheme(region=mypal)

lp <- levelplot(stack_rs, 
          margin=FALSE,
          layer=c('rs_annual_1', 'rs_spring_1','rs_summer_1',
                  'rs_annual_2', 'rs_spring_2','rs_summer_2'),
          par.settings=myTheme,
          pretty=TRUE,
          names.attr=c("Annual 2005", "Spring 2005", "Summer 2005",
                       "Annual 2012", "Summer 2012", "Spring 2012"),
          layout=c(3,2),
          main='Resilience (rs)', xlab=NULL, ylab=NULL)

print(lp)

exportpdf(mypdf=paste0(di, '/images/raster_maps/raster_rs.pdf'), lp) 
```

### Resistance 
```{r, echo=FALSE, fig.cap='Resistance', fig.height=7, fig.width=10}
# Resistance  

# Select a palette http://colorbrewer2.org/
mypal <- brewer.pal(9, "RdYlGn")
# Specify the color palette
myTheme=rasterTheme(region=mypal)

lp <- levelplot(stack_rt, 
                margin=FALSE,
                layer=c('rt_annual_1', 'rt_spring_1','rt_summer_1',
                        'rt_annual_2', 'rt_spring_2','rt_summer_2'),
                par.settings=myTheme,
                pretty=TRUE,
                names.attr=c("Annual 2005", "Spring 2005", "Summer 2005",
                             "Annual 2012", "Summer 2012", "Spring 2012"),
                layout=c(3,2),
                main='Resistance (rt)', xlab=NULL, ylab=NULL)

print(lp)

exportpdf(mypdf=paste0(di, '/images/raster_maps/raster_rt.pdf'), lp) 
```


### Recovery 
```{r, echo=FALSE, fig.cap='Recovery', fig.height=7, fig.width=10}
# Recovery 

# Select a palette http://colorbrewer2.org/
mypal <- brewer.pal(9, "RdYlGn")
# Specify the color palette
myTheme=rasterTheme(region=mypal)

lp <- levelplot(stack_rc, 
                margin=FALSE,
                layer=c('rc_annual_1', 'rc_spring_1','rc_summer_1',
                        'rc_annual_2', 'rc_spring_2','rc_summer_2'),
                par.settings=myTheme,
                pretty=TRUE,
                names.attr=c("Annual 2005", "Spring 2005", "Summer 2005",
                             "Annual 2012", "Summer 2012", "Spring 2012"),
                layout=c(3,2),
                main='Recovery (rc)', xlab=NULL, ylab=NULL)

print(lp)

exportpdf(mypdf=paste0(di, '/images/raster_maps/raster_rc.pdf'), lp) 
```


### Relative Resilience
```{r, echo=FALSE, fig.cap='Relative Resilience', fig.height=7, fig.width=10}
# Relative Resilience 

# Select a palette http://colorbrewer2.org/
mypal <- brewer.pal(9, "RdYlGn")
# Specify the color palette
myTheme=rasterTheme(region=mypal)

lp <- levelplot(stack_rrs, 
                margin=FALSE,
                layer=c('rrs_annual_1', 'rrs_spring_1','rrs_summer_1',
                        'rrs_annual_2', 'rrs_spring_2','rrs_summer_2'),
                par.settings=myTheme,
                pretty=TRUE,
                names.attr=c("Annual 2005", "Spring 2005", "Summer 2005",
                             "Annual 2012", "Summer 2012", "Spring 2012"),
                layout=c(3,2),
                main='Relative Resilience (rs)', xlab=NULL, ylab=NULL)

print(lp)

exportpdf(mypdf=paste0(di, '/images/raster_maps/raster_rrs.pdf'), lp) 
```



# Elevation pattern 

* We remove population `Cadiar`

```{r}
# Prepare data
elev <- read.csv(file=paste(di, "/data/elev.csv", sep=""), header = TRUE, sep = ',')

eviresi <- evi_resilience %>% 
  # join elevation data
  dplyr::inner_join(elev, by='iv_malla_modi_id') %>% 
  # Add a variable for population cluster
  mutate(clu_pop = ifelse(pop %in% c(1,2,3,4,5), 'a', 'b'))


```

## Explore elevation pattern general and by cluster of populations

```{r}
# Change format of the dataset (wide to long)
df_melt <- melt(eviresi, id.vars = c('iv_malla_modi_id', 'pop',
                                      'long', 'lat', 'elev', 'clu_pop', 
                                     'event', 'seasonF'))

df_aux <- df_melt %>% filter(variable %in% c('rs','rt','rc', 'rrs'))

  
label_variable <- c('rt' = 'Resistance', 
                    'rc' = 'Recovery',
                    'rs' = 'Resilience',
                    'rrs' = 'Relative Resilience',
                    '0_pre' = 'PreDrought',
                    '1_dr' = 'Drought',
                    '2_post' = 'PostDrought')
label_cluster <- c('a' = 'Northern slope',
                   'b' = 'Southern slope')

label_season <- c('annual' = 'Annual',
                  'summer' = 'Summer',
                  'spring' = 'Spring')
```

### General pattern 
```{r, fig.cap='Resilience components vs. elevation', fig.width=12, fig.height=8}
g <- df_aux %>% 
  filter(variable %in% c('rt', 'rc', 'rs')) %>% 
  ggplot(aes(x=elev, y=value)) + 
  geom_point(col='gray') + 
  geom_smooth(method = 'lm') + 
  facet_grid(seasonF~variable,
             labeller = labeller(.cols = label_variable,
                                 .rows = label_season)) +
             # labeller = as_labeller(label_variable)) + 
  theme_bw() + 
  theme(strip.background = element_rect(fill = "white")) 

g
pdf(file=paste0(di, "/images/plot_resi_elev.pdf"), height = 9, width = 9)
g
dev.off() 
```


```{r, fig.cap='Relative Resilience vs. elevation'}
gr <- df_aux %>% 
  filter(variable == 'rrs') %>%
  ggplot(aes(x=elev, y=value)) + 
  geom_point(col='gray') + 
  theme_bw() +
  geom_smooth(method = 'lm') + 
  facet_grid(seasonF~.)
  ggtitle('Relative Resilience')

gr

pdf(file=paste0(di, "/images/plot_resi_rel_elev.pdf"), height = 8, width = 5)
gr
dev.off() 
```


### Elevational pattern by population 
#### Annual 

```{r, fig.cap='Resilience components vs. elevation (grouped by cluster) - Annual', fig.width=12, fig.height=10}
gp <- 
  df_aux %>% filter(variable %in% c('rt', 'rc', 'rs')) %>% filter(seasonF == 'annual') %>%  
  ggplot( aes(x=elev, y=value)) +
  geom_point(col='gray') + 
  geom_smooth(method = 'lm') + 
  facet_grid(variable ~ clu_pop, scales = 'free_y',
             labeller = labeller(.rows = label_variable,
                                 .cols = label_cluster)) +
  # facet_wrap(~variable, labeller = as_labeller(label_variable)) + 
  theme_bw() + ggtitle('Annual') +
  theme(strip.background = element_rect(fill = "white")) 
gp 

pdf(file=paste0(di, "/images/plot_resicomp_elev_grouped_annual.pdf"), height = 8, width = 8)
gp
dev.off() 
```

#### Spring
```{r, fig.cap='Resilience components vs. elevation (grouped by cluster) - Spring', fig.width=12, fig.height=10}
gp <- 
  df_aux %>% filter(variable %in% c('rt', 'rc', 'rs')) %>% filter(seasonF == 'spring') %>%  
  ggplot( aes(x=elev, y=value)) +
  geom_point(col='gray') + 
  geom_smooth(method = 'lm') + 
  facet_grid(variable ~ clu_pop, scales = 'free_y',
             labeller = labeller(.rows = label_variable,
                                 .cols = label_cluster)) +
  # facet_wrap(~variable, labeller = as_labeller(label_variable)) + 
  theme_bw() + ggtitle('Spring') +
  theme(strip.background = element_rect(fill = "white")) 
gp 

pdf(file=paste0(di, "/images/plot_resicomp_elev_grouped_spring.pdf"), height = 8, width = 8)
gp
dev.off() 
```

#### Summer
```{r, fig.cap='Resilience components vs. elevation (grouped by cluster) - Spring', fig.width=12, fig.height=10}
gp <- 
  df_aux %>% filter(variable %in% c('rt', 'rc', 'rs')) %>% filter(seasonF == 'summer') %>%  
  ggplot( aes(x=elev, y=value)) +
  geom_point(col='gray') + 
  geom_smooth(method = 'lm') + 
  facet_grid(variable ~ clu_pop, scales = 'free_y',
             labeller = labeller(.rows = label_variable,
                                 .cols = label_cluster)) +
  # facet_wrap(~variable, labeller = as_labeller(label_variable)) + 
  theme_bw() + ggtitle('Summer') +
  theme(strip.background = element_rect(fill = "white")) 
gp 

pdf(file=paste0(di, "/images/plot_resicomp_elev_grouped_summer.pdf"), height = 8, width = 8)
gp
dev.off() 
```




# Explore pattern by population (cluster)
### Event 1 (2005)
```{r, fig.cap='Resilience components by cluster (populations)', fig.width=9, fig.height=9}
gpop_2005 <- df_aux %>% filter(event == 1) %>% 
  ggplot(aes(x=clu_pop, y=value)) + 
  geom_boxplot() + 
  facet_grid(variable~seasonF, scales = 'free_y',
             labeller = labeller(.rows = label_variable,
                                 .cols = label_season)) +
  # facet_wrap(~variable, scales = 'free_y', labeller = as_labeller(label_variable)) + 
  theme_bw() + xlab('')+
  theme(strip.background = element_rect(fill = "white")) +  
  ggtitle('2005') + 
  scale_x_discrete(labels = label_cluster)

gpop_2005
pdf(file=paste0(di, "/images/plot_resicomp_by_cluster_event1.pdf"), height = 7, width = 8)
gpop_2005
dev.off()
```

### Event 2 (2012)
```{r, fig.cap='Resilience components by cluster (populations)', fig.width=9, fig.height=9}
gpop_2012 <- df_aux %>% filter(event == 2) %>% 
  ggplot(aes(x=clu_pop, y=value)) + 
  geom_boxplot() + 
  facet_grid(variable~seasonF, scales = 'free_y',
             labeller = labeller(.rows = label_variable,
                                 .cols = label_season)) +
  # facet_wrap(~variable, scales = 'free_y', labeller = as_labeller(label_variable)) + 
  theme_bw() + xlab('')+
  theme(strip.background = element_rect(fill = "white")) + 
  ggtitle('2012') + 
  scale_x_discrete(labels = label_cluster)

gpop_2012
pdf(file=paste0(di, "/images/plot_resicomp_by_cluster_event2.pdf"), height = 7, width = 8)
gpop_2012
dev.off()
```

# Bar plots
## Raw variables 

```{r, fig.cap='Resilience components by cluster (populations)', fig.width=10, fig.height=4}

# Bar plot 
variables <- c('dr','post', 'pre','rs','rc','rt','rrs')
myseasons <- unique(evi_resilience$seasonF)
auxdf <- data.frame()

for (m in myseasons){ 
  # filter by season
  aux_season <- eviresi %>% filter(seasonF == m)
  
  for (j in c(1:2)){
    # filter by event 
    auxilio <- aux_season %>% filter(event==j) 
    
    for (i in variables){ 
      aux <- auxilio %>% 
        dplyr::group_by(clu_pop) %>% 
        summarise_each_(funs(mean, sd, se=sd(.)/sqrt(n())), i) %>% 
        mutate(variable=i) %>% 
        mutate(event = j) %>% 
        mutate(seasonF = m)
      
      auxdf <- rbind(auxdf, aux)
    }
  }
}

auxdf <- auxdf %>% mutate(variable = plyr::mapvalues(variable,
                                                         c("dr","post","pre","rs","rc","rt","rrs"),
                                                         c("1_dr","2_post","0_pre","rs","rc","rt","rrs")),
                          event = plyr::mapvalues(event, c(1,2), c(2005,2012)))
```

```{r}
col2012 <- '#0700fe'
col2005 <- '#19e00b'


gpop_bar <- auxdf %>% 
  filter(variable %in% c('1_dr', '0_pre', '2_post')) %>% filter(seasonF=='annual') %>% 
  ggplot(aes(x=variable, y=mean, fill=as.factor(event))) + 
  geom_bar(stat='identity', position="dodge") +# fill='black', colour='black') + 
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd, colour=as.factor(event)), 
                width=.2, position=position_dodge(.9))+ 
  facet_wrap(~clu_pop, labeller = as_labeller(label_cluster)) + 
  theme_bw() + xlab('') + 
  theme(strip.background = element_rect(fill = "white")) +
  scale_colour_manual(values=c(col2005, col2012),name='') + 
  scale_fill_manual(values=c(col2005, col2012), name='') +
  ylab('Annual EVI') +
  scale_x_discrete(labels = label_variable) 

gpop_bar
pdf(file=paste0(di, "/images/plot_rawcomp_bar_by_cluster_annual.pdf"), height = 6, width = 8)
gpop_bar
dev.off()
```

```{r}
gpop_bar <- auxdf %>% 
  filter(variable %in% c('1_dr', '0_pre', '2_post')) %>% filter(seasonF=='summer') %>% 
  ggplot(aes(x=variable, y=mean, fill=as.factor(event))) + 
  geom_bar(stat='identity', position="dodge") +# fill='black', colour='black') + 
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd, colour=as.factor(event)), 
                width=.2, position=position_dodge(.9))+ 
  facet_wrap(~clu_pop, labeller = as_labeller(label_cluster)) + 
  theme_bw() + xlab('') + 
  theme(strip.background = element_rect(fill = "white")) +
  scale_colour_manual(values=c(col2005, col2012),name='') + 
  scale_fill_manual(values=c(col2005, col2012), name='') +
  ylab('Summer EVI') +
  scale_x_discrete(labels = label_variable) 

gpop_bar
pdf(file=paste0(di, "/images/plot_rawcomp_bar_by_cluster_summer.pdf"), height = 6, width = 8)
gpop_bar
dev.off()
```

```{r}
gpop_bar <- auxdf %>% 
  filter(variable %in% c('1_dr', '0_pre', '2_post')) %>% filter(seasonF=='spring') %>% 
  ggplot(aes(x=variable, y=mean, fill=as.factor(event))) + 
  geom_bar(stat='identity', position="dodge") +# fill='black', colour='black') + 
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd, colour=as.factor(event)), 
                width=.2, position=position_dodge(.9))+ 
  facet_wrap(~clu_pop, labeller = as_labeller(label_cluster)) + 
  theme_bw() + xlab('') + 
  theme(strip.background = element_rect(fill = "white")) +
  scale_colour_manual(values=c(col2005, col2012),name='') + 
  scale_fill_manual(values=c(col2005, col2012), name='') +
  ylab('Spring EVI') +
  scale_x_discrete(labels = label_variable) 

gpop_bar
pdf(file=paste0(di, "/images/plot_rawcomp_bar_by_cluster_spring.pdf"), height = 6, width = 8)
gpop_bar
dev.off()
```

## Resilience bar 

```{r}
gpop_bar <- auxdf %>% 
  filter(variable %in% c('rs', 'rc', 'rt')) %>% filter(seasonF=='annual') %>% 
  ggplot(aes(x=clu_pop, y=mean, fill=as.factor(event))) + 
  geom_bar(stat='identity', position="dodge") +# fill='black', colour='black') + 
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd, colour=as.factor(event)), 
                width=.2, position=position_dodge(.9))+ 
  facet_wrap(~variable, labeller = as_labeller(label_variable)) + 
  theme_bw() + xlab('') + 
  theme(strip.background = element_rect(fill = "white")) +
  scale_colour_manual(values=c(col2005, col2012),name='') + 
  scale_fill_manual(values=c(col2005, col2012), name='') +
  ylab('Annual EVI') +
  scale_x_discrete(labels = label_cluster) 

gpop_bar
pdf(file=paste0(di, "/images/plot_resicomp_bar_by_cluster_annual.pdf"), height = 6, width = 8)
gpop_bar
dev.off()
```


```{r}
gpop_bar <- auxdf %>% 
  filter(variable %in% c('rs', 'rc', 'rt')) %>% filter(seasonF=='summer') %>% 
  ggplot(aes(x=clu_pop, y=mean, fill=as.factor(event))) + 
  geom_bar(stat='identity', position="dodge") +# fill='black', colour='black') + 
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd, colour=as.factor(event)), 
                width=.2, position=position_dodge(.9))+ 
  facet_wrap(~variable, labeller = as_labeller(label_variable)) + 
  theme_bw() + xlab('') + 
  theme(strip.background = element_rect(fill = "white")) +
  scale_colour_manual(values=c(col2005, col2012),name='') + 
  scale_fill_manual(values=c(col2005, col2012), name='') +
  ylab('Summer EVI') +
  scale_x_discrete(labels = label_cluster) 

gpop_bar
pdf(file=paste0(di, "/images/plot_resicomp_bar_by_cluster_summer.pdf"), height = 6, width = 8)
gpop_bar
dev.off()
```

```{r}
gpop_bar <- auxdf %>% 
  filter(variable %in% c('rs', 'rc', 'rt')) %>% filter(seasonF=='spring') %>% 
  ggplot(aes(x=clu_pop, y=mean, fill=as.factor(event))) + 
  geom_bar(stat='identity', position="dodge") +# fill='black', colour='black') + 
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd, colour=as.factor(event)), 
                width=.2, position=position_dodge(.9))+ 
  facet_wrap(~variable, labeller = as_labeller(label_variable)) + 
  theme_bw() + xlab('') + 
  theme(strip.background = element_rect(fill = "white")) +
  scale_colour_manual(values=c(col2005, col2012),name='') + 
  scale_fill_manual(values=c(col2005, col2012), name='') +
  ylab('Spring EVI') +
  scale_x_discrete(labels = label_cluster) 

gpop_bar
pdf(file=paste0(di, "/images/plot_resicomp_bar_by_cluster_spring.pdf"), height = 6, width = 8)
gpop_bar
dev.off()
```


# Exploring relationships 

```{r, echo=FALSE, eval=FALSE}
# https://cran.r-project.org/web/packages/broom/vignettes/broom_and_dplyr.html

# Regresssion models resilience component ~ elevation by cluster pop 

# 'rs' 
myformula <- as.formula(rs ~ elev) 

# Get coef
reg_coef_rs <- eviresi_f %>% dplyr::group_by(clu_pop) %>%
    do(fit = lm(myformula, data =.)) %>% broom::tidy(fit) %>% mutate(variable = 'rs')
# Get stat regression
reg_summ_rs <- eviresi_f %>% dplyr::group_by(clu_pop) %>%
    do(fit = lm(myformula, data =.)) %>%  broom::glance(fit) %>% mutate(variable = 'rs')


# 'rt' 
myformula <- as.formula(rt ~ elev) 

# Get coef
reg_coef_rt <- eviresi_f %>% dplyr::group_by(clu_pop) %>%
    do(fit = lm(myformula, data =.)) %>% broom::tidy(fit) %>% mutate(variable = 'rt')
# Get stat regression
reg_summ_rt <- eviresi_f %>% dplyr::group_by(clu_pop) %>%
    do(fit = lm(myformula, data =.)) %>%  broom::glance(fit) %>% mutate(variable = 'rt')


# 'rc' 
myformula <- as.formula(rc ~ elev) 

# Get coef
reg_coef_rc <- eviresi_f %>% dplyr::group_by(clu_pop) %>%
    do(fit = lm(myformula, data =.)) %>% broom::tidy(fit) %>% mutate(variable = 'rc')
# Get stat regression
reg_summ_rc <- eviresi_f %>% dplyr::group_by(clu_pop) %>%
    do(fit = lm(myformula, data =.)) %>%  broom::glance(fit) %>% mutate(variable = 'rc')


# 'rrs' 
myformula <- as.formula(rrs ~ elev) 

# Get coef
reg_coef_rrs <- eviresi_f %>% dplyr::group_by(clu_pop) %>%
    do(fit = lm(myformula, data =.)) %>% broom::tidy(fit) %>% mutate(variable = 'rrs')
# Get stat regression
reg_summ_rrs <- eviresi_f %>% dplyr::group_by(clu_pop) %>%
    do(fit = lm(myformula, data =.)) %>%  broom::glance(fit) %>% mutate(variable = 'rrs')



# Create table with all regressions output 

reg_coef <- rbind(reg_coef_rs, reg_coef_rt, reg_coef_rc, reg_coef_rrs)
reg_summ <- rbind(reg_summ_rs, reg_summ_rt, reg_summ_rc, reg_summ_rrs)
```

```{r}
kable(reg_coef, caption = "Coefficients of Regression")

kable(reg_summ[,c(1:3,6,13)], caption = "Coefficients of Regression")
```



```{r, echo=FALSE, eval=FALSE} 
m <- 
  eviresi_f %>% 
  split(.$clu_pop) %>% 
  map(~ lm(rs ~ elev, data =.)) %>% 
  map_df(broom::tidy, .id="i") 






```







