---
title: "Explore anomalies"
author: "AJ Perez-Luque (@ajpelu)"
date: "2017 Jan"
output:  
  md_document:
    variant: markdown_github
---
```{r, echo=FALSE, message=FALSE}
require(knitr)
opts_chunk$set(fig.align='center', fig.width = 14, message = FALSE, warning = FALSE) 
```

```{r wd, echo=FALSE}
#---------------------------------
machine <- 'ajpelu'
# machine <- 'ajpeluLap'
di <- paste('/Users/', machine, '/Dropbox/phd/phd_repos/qpyr_resilience', sep='')
#---------------------------------
```


```{r packages, warning=FALSE, message=FALSE}
library("rgdal")
library("sp")
library("raster")
library("rasterVis")
source(paste0(di,"/R/exportpdf.R")) # function to export raster levelplots maps as pdf
library("dplyr")
library("ggplot2")
library("reshape2")
library("purrr")
# library("pander")
library("knitr")
```

## Read data and Prepare Data 
```{r}
# Read data 
anomalias_season <- read.csv(file=paste(di, "/data/anomalies/anomalias_season.csv", sep=""), header = TRUE, sep = ',')
anomalias_composite <- read.csv(file=paste(di, "/data/anomalies/anomalias_composite.csv", sep=""), header = TRUE, sep = ',')


anomalias_season <- anomalias_season %>% 
  mutate(clu_pop = as.factor(ifelse(pop == 1, 'Camarate',
                             ifelse(pop %in% c(2,3,4,5), 'Northern slope',
                             ifelse(pop %in% c(6,7,8), 'Southern slope', 'out'))))) %>% 
    filter(clu_pop != 'out') 


anomalias_composite <- anomalias_composite %>% 
  mutate(clu_pop = as.factor(ifelse(pop == 1, 'Camarate',
                             ifelse(pop %in% c(2,3,4,5), 'Northern slope',
                             ifelse(pop %in% c(6,7,8), 'Southern slope', 'out'))))) %>% 
    filter(clu_pop != 'out')
```

# Season data 
## Get mean values of anomalies by clu_pop
```{r}
summ_aS <- anomalias_season %>% 
  # filter(y != 2000) %>% 
  group_by(clu_pop, y, composite) %>% 
  summarise(mean=mean(anomaly),
            sd = sd(anomaly),
            se = sd/sqrt(length(anomaly))) %>% 
 mutate(signo = ifelse(mean >= 0, 'pos', 'neg'))
               

summ_asS <- anomalias_season %>% 
  filter(y != 2000) %>%
  group_by(clu_pop, y, composite) %>% 
  summarise(mean=mean(anomaly_std),
            sd = sd(anomaly_std),
            se = sd/sqrt(length(anomaly_std))) %>% 
  mutate(signo = ifelse(mean >= 0, 'pos', 'neg'))
```


```{r, echo=FALSE}
# colours 
# color_neg <- '#d73027'
color_neg <- '#a63603'
# darkblue
#color_pos <- '#1a9850'
color_pos <- '#006d2c'
```

## Plot anomaly 
### All seasons 

```{r}
anomalo_season_all_cluster_pop <- ggplot(summ_aS, aes(x=y, y=mean, fill=signo)) + 
  geom_bar(stat = "identity") + 
  geom_errorbar(aes(ymin = mean - 2*se, ymax = mean + 2*se, colour=signo), width=.5) +
  facet_grid(clu_pop ~ composite) + 
  scale_fill_manual(values = c("pos" = color_pos, "neg" =  color_neg)) +
  scale_color_manual(values = c("pos" = color_pos, "neg" =  color_neg)) +
  ylab('Anomaly') + xlab('') +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        strip.background = element_rect(fill = "white"))


#pdf(file=paste0(di, "/images/anomalo_season_all_cluster_pop.pdf"), height = 10, width = 16)
anomalo_season_all_cluster_pop
#dev.off()
```

## Only Summer and Spring

```{r}
anomalo_season_su_sp_cluster_pop <- summ_aS %>% filter(composite %in% c('spring', 'summer')) %>% 
ggplot(aes(x=y, y=mean, fill=signo)) + 
  geom_bar(stat = "identity") + 
  geom_errorbar(aes(ymin = mean - 2*se, ymax = mean + 2*se, colour=signo), width=.5) +
  facet_grid(clu_pop ~ composite) + 
  scale_fill_manual(values = c("pos" = color_pos, "neg" =  color_neg)) +
  scale_color_manual(values = c("pos" = color_pos, "neg" =  color_neg)) +
  ylab('Anomaly') + xlab('') +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        strip.background = element_rect(fill = "white"))

#pdf(file=paste0(di, "/images/anomalo_season_su_sp_cluster_pop.pdf"), height = 6, width = 10)
anomalo_season_su_sp_cluster_pop
# dev.off()
```

## Plot anomaly standardized 
### All seasons 
```{r}
ggplot(summ_asS, aes(x=y, y=mean, fill=signo)) + 
  geom_bar(stat = "identity") + 
  geom_errorbar(aes(ymin = mean - 2*se, ymax = mean + 2*se, colour=signo), width=.5) +
  facet_grid(clu_pop ~ composite) + 
  scale_fill_manual(values = c("pos" = color_pos, "neg" =  color_neg)) +
  scale_color_manual(values = c("pos" = color_pos, "neg" =  color_neg)) +
  ylab('Standardized anomaly') + xlab('') +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        strip.background = element_rect(fill = "white"))
```

## Only Summer and Spring

```{r}
summ_asS %>% filter(composite %in% c('spring', 'summer')) %>% 
ggplot(aes(x=y, y=mean, fill=signo)) + 
  geom_bar(stat = "identity") + 
  geom_errorbar(aes(ymin = mean - 2*se, ymax = mean + 2*se, colour=signo), width=.5) +
  facet_grid(clu_pop ~ composite) + 
  scale_fill_manual(values = c("pos" = color_pos, "neg" =  color_neg)) +
  scale_color_manual(values = c("pos" = color_pos, "neg" =  color_neg)) +
  ylab('Standardized anomaly') + xlab('') +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        strip.background = element_rect(fill = "white"))
```


### Combine N and S (not Camarate)
```{r}
anomalias_seasonNS <- anomalias_season %>% 
  mutate(clu_pop = as.factor(ifelse(pop %in% c(1,2,3,4,5), 'Northern slope',
                             ifelse(pop %in% c(6,7,8), 'Southern slope', 'out')))) %>% 
    filter(clu_pop != 'out') 

summNS_aS <- anomalias_seasonNS %>% 
  # filter(y != 2000) %>% 
  group_by(clu_pop, y, composite) %>% 
  summarise(mean=mean(anomaly),
            sd = sd(anomaly),
            se = sd/sqrt(length(anomaly))) %>% 
 mutate(signo = ifelse(mean >= 0, 'pos', 'neg'))
               

summNS_asS <- anomalias_seasonNS %>% 
  filter(y != 2000) %>%
  group_by(clu_pop, y, composite) %>% 
  summarise(mean=mean(anomaly_std),
            sd = sd(anomaly_std),
            se = sd/sqrt(length(anomaly_std))) %>% 
  mutate(signo = ifelse(mean >= 0, 'pos', 'neg'))

```


## Plot anomaly 
### All seasons 

```{r}
anomalo_season_all_cluster_popNS <- ggplot(summNS_aS, aes(x=y, y=mean, fill=signo)) + 
  geom_bar(stat = "identity") + 
  geom_errorbar(aes(ymin = mean - 2*se, ymax = mean + 2*se, colour=signo), width=.5) +
  facet_grid(clu_pop ~ composite) + 
  scale_fill_manual(values = c("pos" = color_pos, "neg" =  color_neg)) +
  scale_color_manual(values = c("pos" = color_pos, "neg" =  color_neg)) +
  ylab('Anomaly') + xlab('') +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        strip.background = element_rect(fill = "white"))


#pdf(file=paste0(di, "/images/anomalo_season_all_cluster_popNS.pdf"), height = 10, width = 16)
anomalo_season_all_cluster_popNS
#dev.off()
```

## Only Summer and Spring

```{r}
anomalo_season_su_sp_cluster_popNS <- summNS_aS %>% filter(composite %in% c('spring', 'summer')) %>% 
ggplot(aes(x=y, y=mean, fill=signo)) + 
  geom_bar(stat = "identity") + 
  geom_errorbar(aes(ymin = mean - 2*se, ymax = mean + 2*se, colour=signo), width=.5) +
  facet_grid(clu_pop ~ composite) + 
  scale_fill_manual(values = c("pos" = color_pos, "neg" =  color_neg)) +
  scale_color_manual(values = c("pos" = color_pos, "neg" =  color_neg)) +
  ylab('Anomaly') + xlab('') +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        strip.background = element_rect(fill = "white"))

# pdf(file=paste0(di, "/images/anomalo_season_su_sp_cluster_popNS.pdf"), height = 6, width = 10)
anomalo_season_su_sp_cluster_popNS
# dev.off()
``` 

# Maps

* Create raster maps (two projections: `epsg:4326` and `epsg:23030`)

```{r, echo=FALSE}

# Create a spatial dataframe 
anom_sp <- SpatialPointsDataFrame(coords = anomalias_season[,c('long','lat')], data = anomalias_season,
                               proj4string = CRS("+init=epsg:4326"))

# Transform 
anom_sp <- spTransform(anom_sp, CRS("+init=epsg:23030"))

myyear <- c(2005,2012) 
seasons <- unique(anomalias_season$composite)

for (j in myyear){ 
  # Filter by year 
  aux_spatial_year <- anom_sp[anom_sp$y ==j,]
  
  for (i in seasons){ 
    aux_spatial <- aux_spatial_year[aux_spatial_year$composite == i,]
    
    # raster auxiliar 
    aux_rast <- raster(aux_spatial, resolution=250)
    
    # raster_layer 
    r_season <- rasterize(aux_spatial, aux_rast, 'anomaly', fun=mean)
    names(r_season) <- paste0('a_',i,'_', j)
    
    # reprojected raster
    r_season_re <- projectRaster(r_season, crs=crs("+init=epsg:4326"))
    names(r_season_re) <- paste0('a_',i,'_',j)
    
    # assign 
    name_raster <- paste0('r_anomalia_', i, '_', j) 
    name_raster_re <- paste0('r_anomalia_', i, '_', j, "_re") 
    
    assign(name_raster, r_season)
    assign(name_raster_re, r_season_re)
    
    writeRaster(r_season, file=paste(di, "/data/raster/r_anomalia_", i, "_", j, ".asc", sep=""), overwrite=TRUE)
    writeRaster(r_season_re, file=paste(di, "/data/raster/r_anomalia_", i, "_", j, "_re", ".tiff", sep=""), overwrite=TRUE) 
  }
}
```



```{r}
# Create stack of raster 
stack_a <- stack(r_anomalia_annual_2005, r_anomalia_spring_2005, r_anomalia_summer_2005, r_anomalia_autumn_2005,
             r_anomalia_winter_2005, r_anomalia_annual_2012, r_anomalia_spring_2012, r_anomalia_summer_2012,
             r_anomalia_autumn_2012, r_anomalia_winter_2012)

stack_a_re <- stack(r_anomalia_annual_2005_re, r_anomalia_spring_2005_re, r_anomalia_summer_2005_re,
                    r_anomalia_autumn_2005_re, r_anomalia_winter_2005_re, r_anomalia_annual_2012_re,
                    r_anomalia_spring_2012_re, r_anomalia_summer_2012_re,
                    r_anomalia_autumn_2012_re, r_anomalia_winter_2012_re)

# Export stack 
temp <- getwd()
setwd(paste(di, "/data/raster/", sep=""))
writeRaster(stack_a, filename = 'r_a_stack', overwrite =TRUE) 
writeRaster(stack_a_re, filename = 'r_a_re_stack', overwrite =TRUE) 
setwd(temp)
```


## Spatial exploration of the anomaly

### Annual 
```{r}
# Select a palette http://colorbrewer2.org/
mypal <- brewer.pal(11, "RdYlGn")
# Specify the color palette
myTheme=rasterTheme(region=mypal)

lp <- levelplot(stack_a_re, 
          margin=FALSE,
          layer=c('a_annual_2005', 'a_annual_2012'),
          #par.settings=myTheme, 
          par.settings=RdBuTheme, 
          names.attr=c("2005", "2012"),
          pretty=TRUE,
          layout=c(2,1),
          main='EVI Annual Anomaly', xlab=NULL, ylab=NULL) 

print(lp)

tiff(filename=paste0(di, '/images/raster_maps/a_2005_2012_annual.tiff'), 
     height = 1600, width = 2500, res=150, pointsize = 20, bg='transparent')
print(lp)
dev.off()

exportpdf(mypdf=paste0(di, '/images/raster_maps/a_2005_2012_annual.pdf'), lp) 
```


### Spring 
```{r}
# Select a palette http://colorbrewer2.org/
mypal <- brewer.pal(11, "RdYlGn")
# Specify the color palette
myTheme=rasterTheme(region=mypal)

lp <- levelplot(stack_a_re, 
          margin=FALSE,
          layer=c('a_spring_2005', 'a_spring_2012'),
          #par.settings=myTheme, 
          par.settings=RdBuTheme, 
          names.attr=c("2005", "2012"),
          pretty=TRUE,
          layout=c(2,1),
          main='EVI Spring Anomaly', xlab=NULL, ylab=NULL) 

print(lp)

tiff(filename=paste0(di, '/images/raster_maps/a_2005_2012_spring.tiff'), 
     height = 1600, width = 2500, res=150, pointsize = 20, bg='transparent')
print(lp)
dev.off()

exportpdf(mypdf=paste0(di, '/images/raster_maps/a_2005_2012_spring.pdf'), lp) 
```


### Summer 
```{r}
# Select a palette http://colorbrewer2.org/
mypal <- brewer.pal(11, "RdYlGn")
# Specify the color palette
myTheme=rasterTheme(region=mypal)

lp <- levelplot(stack_a_re, 
          margin=FALSE,
          layer=c('a_summer_2005', 'a_summer_2012'),
          #par.settings=myTheme, 
          par.settings=RdBuTheme, 
          names.attr=c("2005", "2012"),
          pretty=TRUE,
          layout=c(2,1),
          main='EVI Summer Anomaly', xlab=NULL, ylab=NULL) 

print(lp)

tiff(filename=paste0(di, '/images/raster_maps/a_2005_2012_summer.tiff'), 
     height = 1600, width = 2500, res=150, pointsize = 20, bg='transparent')
print(lp)
dev.off()

exportpdf(mypdf=paste0(di, '/images/raster_maps/a_2005_2012_summer.pdf'), lp) 
```

