---
title: "prepare data of EVI for resilience computation"
author: "AJ Perez-Luque (@ajpelu)"
date: "2016 May"
output:  
  md_document:
    variant: markdown_github
---
  
  
```{r wd, echo=FALSE}
#---------------------------------
machine <- 'ajpelu'
# machine <- 'ajpeluLap'
di <- paste('/Users/', machine, '/Dropbox/phd/phd_repos/qpyr_resilience', sep='')
#---------------------------------
```

```{r packages, warning=FALSE}
library("dplyr")
library("lubridate")
library("ggplot2")
library("knitr")
library('printr')
source(paste0(di,"/R/getComposite.R"))
```

## Read and prepare evi data

We created three datasets: 

* annual and seasonal evi by pixel (output as `./data/evi_atributes_all.csv`)
* annual and seasonal ndvi by pixel (output as `./data/ndvi_atributes_all.csv`)
* iv by pixel and by composite (output as `./data/iv_composite.csv`)

The first two dataframes have the following fields: 

* `iv_malla_modi_id`: the identifier of the modis cell
* `year`
* `evi` or `ndvi`: the value of the EVI (or NDVI) (cumulative value for each season)
* `season`: the season of cumulative evi:
  * `0` annual value
  * `1` spring value
  * `2` summer value 
  * `3` autumn value 
  * `4` winter value 
* `seasonF`: the season coded as factor 
* `long`: longitude coordinates
* `lat`: latitute coordinates 
* `pop`: numeric code of the *Q. pyrenaica* population 

The iv_composite dataframe has the following fields: 

* `evi` or `ndvi`: value of IV for the date 
* `date`: date of adquisition of the image
* `composite`: number of composite (23 by year)
* `iv_malla_modi_id`, `year`, `lat`, `long`, `pop`, `seasonF` 

```{r}
# Read data
# rawdata <- read.csv(file=paste(di, "/data_raw/evi/iv_quercus_pyrenaica.csv", sep= ""), header = TRUE, sep = ',')
rawdata <- read.csv(file=paste(di, "/data_raw/evi/iv_raw_2016.csv", sep= ""), header = TRUE, sep = ',')
```

## Some metadata of the time series 
### Temporal range of the Time series 
```{r}
# Get temporal range of the data
# Start date
unique(min(as.Date(rawdata$date)))

# End date 
unique(max(as.Date(rawdata$date)))
```

### QA
```{r}
# See n of images per year and per pixel 
n_images_pixel <- rawdata %>% 
  mutate(year = lubridate::year(date)) %>% 
  group_by(year) %>%
  summarise(n = n(),
            n_pixel = n()/length(unique(iv_malla_modi_id)))

kable(n_images_pixel)
```


## Prepare data 
### Get the composite of the images and the season
* See [Testa et al. 2014](https://www.researchgate.net/publication/262566793_Correcting_MODIS_16-day_composite_NDVI_time-series_with_actual_acquisition_dates)
* Use a [custom function](/R/getComposite.R)

```{r}
# Get leap years 
years <- unique(rawdata$year)
ly <- years[leap_year(years)]

# Two functions 
rd <- rawdata %>% 
  mutate(m= lubridate::month(date),
         d= lubridate::day(date)) %>% 
  mutate(composite = ifelse(year %in% ly, 
                            getComposite_leap(m,d),
                            getComposite_nonleap(m,d))) %>% 
  mutate(season = ifelse(composite < 6, 'winter',
                   ifelse(composite < 12, 'spring',
                    ifelse(composite < 18, 'summer','autumn'))))
```

### Scale factor of the NDVI and EVI data 
```{r}
# Apply scale factor https://lpdaac.usgs.gov/dataset_discovery/modis/modis_products_table/mod13q1 
rd <- rd %>% 
  mutate(evi = evi * 0.0001,
         ndvi = ndvi * 0.0001) 
```

### Create dataframe with composites
```{r}
iv_composite <- rd %>% 
  dplyr::select(iv_malla_modi_id, evi, ndvi,pop, date, year, long, lat, composite, seasonF = season)
```

### Create seasonal dataframes of EVI and NDVI 
* EVI
```{r}
# Create annual evi by pixel 
evi_annual <- rd %>% 
  group_by(iv_malla_modi_id, year) %>%
  summarise(evi = sum(evi[evi >=0])) %>%
  mutate(seasonF='annual', 
         season = 0)

# Create seasonal evi by pixel 
evi_season <- rd %>% 
  group_by(iv_malla_modi_id, year, season) %>%
  summarise(evi = sum(evi[evi >=0])) %>% 
  mutate(seasonF = season) %>% 
  mutate(season = ifelse(season == 'autumn', 3, 
                   ifelse(season == 'winter', 4, 
                    ifelse(season == 'spring', 1,2))))

evidf <- rbind(evi_annual, evi_season)

# Add coordinates and pob 
aux_rd <- rd %>% dplyr::select(iv_malla_modi_id, long, lat, pop) %>%
  group_by(iv_malla_modi_id) %>% unique()


# Join dataframes 
evidf <- evidf %>% dplyr::inner_join(aux_rd, by="iv_malla_modi_id") 
```

* NDVI 
```{r}
# Create annual ndvi by pixel 
ndvi_annual <- rd %>% 
  group_by(iv_malla_modi_id, year) %>%
  summarise(ndvi = sum(ndvi[ndvi >=0])) %>%
  mutate(seasonF='annual', 
         season = 0)

# Create seasonal ndvi by pixel 
ndvi_season <- rd %>% 
  group_by(iv_malla_modi_id, year, season) %>%
  summarise(ndvi = sum(ndvi[ndvi >=0])) %>% 
  mutate(seasonF = season) %>% 
  mutate(season = ifelse(season == 'autumn', 3, 
                   ifelse(season == 'winter', 4, 
                    ifelse(season == 'spring', 1,2))))

ndvidf <- rbind(ndvi_annual, ndvi_season)

# Join dataframes 
ndvidf <- ndvidf %>% dplyr::inner_join(aux_rd, by="iv_malla_modi_id")
```

### Export dataframes
```{r}
# Export dataframes 
write.csv(evidf, file=paste(di, "/data/evi_atributes_all.csv", sep=""), row.names = FALSE)
write.csv(ndvidf, file=paste(di, "/data/ndvi_atributes_all.csv", sep=""), row.names = FALSE)
write.csv(iv_composite, file=paste(di, "/data/iv_composite.csv", sep=""), row.names = FALSE)
```
