---
title: "prepare data of EVI for resilience computation"
author: "AJ Perez-Luque (@ajpelu)"
date: "2016 May"
output:  
  md_document:
    variant: markdown_github
---
  
  
```{r wd, echo=FALSE}
#---------------------------------
machine <- 'ajpelu'
# machine <- 'ajpeluLap'
di <- paste('/Users/', machine, '/Dropbox/phd/phd_repos/qpyr_resilience', sep='')
#---------------------------------
```

```{r packages, warning=FALSE}
library("dplyr")
library("lubridate")
library("ggplot2")
```

## Read and prepare evi data

We created two datasets: 

* annual evi by pixel
* seasonal evi by pixel 

Each dataframe has the following fields: 

* `iv_malla_modi_id`: the identifier of the modis cell
* `year`
* `evi_mean`: the value of the evi (cumulative value for each season)
* `season`: the season of cumulative evi:
  * `0` annual value
  * `1` spring value
  * `2` summer value 
  * `3` autumn value 
  * `4` winter value 
* `seasonF`: the season coded as factor 
* `lng`: longitude coordinates
* `lat`: latitute coordinates 
* `poblacion`: numeric code of the *Q. pyrenaica* population 
  

```{r}
# Read and prepare data
rawdata <- read.csv(file=paste(di, "/data_raw/evi/iv_quercus_pyrenaica.csv", sep= ""), header = TRUE, sep = ',')

# Get temporal range of the data
# Start date
unique(min(as.Date(rawdata$date)))

# End date 
unique(max(as.Date(rawdata$date)))

# Create variables of year and month 
# Apply scale factor https://lpdaac.usgs.gov/dataset_discovery/modis/modis_products_table/mod13q1 
rawdata <- rawdata %>% 
  mutate(myevi = evi * 0.0001,
         year = lubridate::year(date),
         month = lubridate::month(date))

# Create annual evi by pixel 
eviyear <- rawdata %>% 
  group_by(iv_malla_modi_id, year) %>%
  summarise(evi = sum(myevi[myevi >=0])) %>%
  mutate(season=0)


# Create seasonal evi 

# Julian date 
# >81 <=173 --> spring (1)
# >173 <=265 --> summer (2)
# > 265 <=356 --> autum (3)
# > 356 and < 81 --> winter (4)
# Get julian day of limits of the season
sp <-lubridate::yday(as.Date("2000-03-21"))
su <- lubridate::yday(as.Date("2000-06-21"))
au <-  lubridate::yday(as.Date("2000-09-21"))
wi <- lubridate::yday(as.Date("2000-12-21"))
season_julian <- c(sp,su,au,wi)

eviseason <- rawdata %>%
  mutate(jday=lubridate::yday(date)) %>%
  select(iv_malla_modi_id, year, myevi, jday) %>% 
  mutate(season = ifelse(jday > 81 & jday <= 173, 1,
                        ifelse(jday > 173 & jday <= 265, 2, 
                               ifelse(jday > 265 & jday <= 356, 3, 4)))) %>%
  group_by(iv_malla_modi_id, year, season) %>%
  summarise(evi = sum(myevi[myevi >=0]))


evidf <- rbind(eviyear, eviseason)

evidf <- evidf %>% 
  mutate(seasonF = ifelse (season == 0, 'annual',
                           ifelse(season == 1, 'spring',
                                  ifelse(season == 2, 'summer',
                                         ifelse(season == 3, 'autumn', 'winter')))))



# Add coordinates and pob 
evi_aux <- rawdata %>% select(iv_malla_modi_id, lng, lat, poblacion) %>%
  group_by(iv_malla_modi_id) %>% unique()


# Join dataframes 
evi <- evidf %>% inner_join(evi_aux, by="iv_malla_modi_id") 

# Export evi dataframe
write.csv(evi, file=paste(di, "/data/evi_attributes_all.csv", sep=""), row.names = FALSE)
```
